---
title: "07_MFA"
author: "Heike Rolker"
date: "2024-01-19"
output: html_document
---

Final file runs in Chapter write-up.


```{r set-up}
# if(!require(devtools))
#   install.packages("devtools", repos = "http://cran.rstudio.com")
# if(!require(thesisdown))
#   devtools::install_github("ismayc/thesisdown")
# library(thesisdown)
# if(!require(kableExtra))
#   devtools::install_github("haozhu233/kableExtra")
# library(kableExtra)
# if(!require(tinytex))
#   devtools::install_github('yihui/tinytex')
# library(tinytex)
# options(tinytex.verbose = TRUE)
# if(!require(flextable))
#   devtools::install_github("davidgohel/flextable")
# library(flextable)
# if(!require(dplyr))
#   install.packages("dplyr", repos = "http://cran.rstudio.com")
# library(dplyr)
# if(!require(ggplot2))
#   install.packages("ggplot2", repos = "http://cran.rstudio.com")
# library(ggplot2)
# if(!require(knitr))
#   install.packages("knitr", repos = "http://cran.rstudio.com")
# library(knitr)
# 
# library(easypackages)
# libraries("rmarkdown", "tidyverse", "pander", "bookdown", "jpeg", "PRISMAstatement", "float", "pacman", "readxl", "kableExtra")
library(tidyverse)
library(data.table)
library(Hmisc)
library(gt)
library(consort)
library(stringr)
library(viridis)
library(hrbrthemes)
library(dplyr)
library(ggplot2)
library(alluvial)
library(ggalluvial)
library(RColorBrewer)
library(plotly)
library(haven)
```
# Housekeeping

```{r functions} 
clear.labels <- (function(x) {
  if(is.list(x)) {
    for(i in seq_along(x)) {
      class(x[[i]]) <- setdiff(class(x[[i]]), 'labelled')
      attr(x[[i]],"label") <- NULL
    }
  } else {
    class(x) <- setdiff(class(x), "labelled")
    attr(x, "label") <- NULL
  }
  return(x)
})


sm <- function(x) sum(x, na.rm = TRUE)
rd <- function(df) df %>% dplyr::mutate_if(is.numeric, round, 1)
md <- function(x) median(x, na.rm=TRUE)
mn <- function(x) mean(x, na.rm=TRUE)

# formatting & other 
sct <- function(x) x/1000 # scale to tonne
scm <- function(x) x/1000000 # scale million (mostly for population)
sc_mt <- function(x) x/1000000000 # scale to million metric ton
mio <- 1000000
ht <- 100000
tt <- 10000
t <- 1000

tblfmt <- function(df){
  df %>% 
    opt_align_table_header(align = "left") %>%
    tab_options(
      table.font.size = px(13)
    )
}
```

Overview:

File to be run before knitting Chapter 3a, data only applicable for this chapter.


Steps for generating MFA/Sankey diagram:


1) estimate missing values, i.e., conduct mass-balance analysis -> in summary file

2) remove excluded households & apply weighting factors & complete lists with non producers - x

3) find solution for NA in type/item when completing missing households

4) re-organise each file as source-destination

# 1) Remove excluded & weigh data for MFA

```{r}
# removal of excluded households -> not required if I weigh!
excl_3a <- read_csv("2_data/final/excl_3a.csv", 
    col_types = cols(...1 = col_skip())) %>% setDT() # should not need to skip, change write function at source

rm_ex <- function(df){
  setDT(df)
  df[!excl_3a, on = .(y4_hhid)]
}

# weighting the data
weighting <- readRDS("2_data/final/weighting.RDS") %>% 
  clear.labels()

weighting %>% inner_join(excl_3a, by = "y4_hhid") # no intersection as it should be!

weigh <- function(df){
 df %>% 
   right_join(weighting, by = "y4_hhid") %>% # only keep households listed in weighting (i.e., those included)
   mutate_if(is.numeric, ~replace(., is.na(.), 0)) %>%
   select(!c(clusterid)) 
}

# previous weighting function
# weigh <- function(df){
#   df <- weighting[df, on=.(y4_hhid)]
#   df <- df %>% mutate(across(is.numeric & !contains("weight_adj"), ~. * weight_adj))
# }

sm <- function(x) sum(x, na.rm = TRUE)
rd <- function(df) df %>% dplyr::mutate_if(is.numeric, round, 1)
md <- function(x) median(x, na.rm=TRUE)
mn <- function(x) mean(x, na.rm=TRUE)
```


```{r mfa data}
# Weigh data, remove excluded households and fill 0 values for non producers (relevant for animal products only)
mfa_crops <- readRDS("2_data/final/mfa_crops.RDS") %>% 
  weigh() %>% 
  mutate(across(c(produced:unallocated), ~. * weight_adj)) # adjust after balancing

mfa_animals <- readRDS("2_data/final/mfa_animals.RDS") %>% 
  ungroup() %>% 
  weigh() %>% 
  mutate(across(c(need:unallocated), ~. * weight_adj))

mfa_eggs <- readRDS("2_data/final/mfa_eggs.RDS") %>% 
  weigh() %>% 
  mutate(across(c(feed:unallocated), ~. * weight_adj))

mfa_milk <- readRDS("2_data/final/mfa_milk.RDS") %>% 
  weigh() %>% 
  mutate(across(c(produced:unallocated), ~. * weight_adj))

mfa_hides <- readRDS("2_data/final/mfa_hides.RDS") %>% 
  weigh() %>% 
  mutate(across(c(slaughter:sold2), ~. * weight_adj))
```

# 2) Uncertainty

Estimates of uncertainty should use unweighted data as it'll be expressed in %?

```{r}

```


# Over and underestimates of allocation to reported production

For crops and milk it is possible to estimate how well allocation and production are in relation to another. For others it may be less meaningful as too few flows are recorded (eggs). For slaughter it is pointless.

Initial approach to generating an uncertainty estimate, will probably have to go through all files again and identify where this can be done and generate a mean where possible.

*Milk*

```{r}
# use unweighted data
milk <- readRDS("2_data/final/mfa_milk.RDS") 

range_milk_fig <- milk %>% 
    weigh() %>% # only included households
  mutate(
    unc = ifelse(uncertain<0, uncertain * -1, uncertain), # should result in positive value
    p.unc = unc/produced,
    p.uncspread = ifelse(uncertain<0, p.unc*-1, p.unc)
  ) %>% 
  select(y4_hhid, type, unc, p.unc, p.uncspread) %>%
  filter(p.unc!=0) %>% 
  dplyr::rename(value= p.uncspread, text = type) %>% 
  ggplot( aes(x=value, color=text, fill=text)) +
    geom_histogram() +
    facet_wrap(~text)

range_milk_fig
```



*Crops*

Harvest as reported and with 2 new missing vars.

```{r eval=FALSE}
# use unweighted data
crops <- readRDS("2_data/final/mfa_crops.RDS") 

range_crops_data <- crops %>% 
  weigh() %>% 
  mutate(
    unc = ifelse(uncertain<0, uncertain * -1, uncertain), # should result in positive value
    p.unc = unc/produced,
    p.uncspread = ifelse(uncertain<0, p.unc*-1, p.unc)
  ) %>% 
  select(y4_hhid, type, unc, p.unc, p.uncspread) %>%
  filter(p.unc!=0) %>% 
  dplyr::rename(value= p.uncspread, text = type) 

write_csv(range_crops_data, "2_data/results/range_crops_data.csv")

range_crops_fig <- range_crops_data %>% 
  ggplot( aes(x=value, color=text, fill=text)) +
    geom_histogram() +
    facet_wrap(~text)

range_crops_fig
```

# Estimate uncertainty

Milk is estimated from average, highest and lowest month, additionally 2 new vars with negative and positive missing.

```{r, eval=FALSE}
uncertainty_milk_stats <- milk %>% 
  mutate(
    upper = ifelse(uncertain<0, uncertain*-1+produced, produced), # production insufficient to cover all disposition
    lower = ifelse(uncertain>0, produced-uncertain, produced) # production lower than all disposition
  ) %>% 
  group_by(type) %>% 
  summarise(
    supper = sum(milk_hi),
    slower = sum(milk_lo),
    supper2 = sum(upper),
    slower2 = sum(lower),
    sum = sum(produced)
  )

uncertainty_milk_stats <- transform(uncertainty_milk_stats, row_stdev=apply(uncertainty_milk_stats, 1, sd, na.rm=TRUE))

uncertainty_milk_stats$Mean <- rowMeans(uncertainty_milk_stats[,2:5])

uncertainty_milk_stats

write_csv(uncertainty_milk_stats, "2_data/results/uncertainty_milk_stats.csv")
```


```{r eval=FALSE}
# 
uncertainty_crops_stats <- crops %>% 
  mutate(
    upper = ifelse(uncertain<0, uncertain*-1+produced, produced), # production insufficient to cover all disposition
    lower = ifelse(uncertain>0, produced-uncertain, produced),# production exceeds disposition, we assume production is lower than reported (this can never be true as the list of uses asked about is limited)
    upper2 = ifelse(uncertain2<0, uncertain2*-1+produced, produced), # same as above, assume processing is a separate entity
    lower2 = ifelse(uncertain2>0, produced-uncertain2, produced),
    meanval = (upper+lower+upper2+lower2+produced)/5) %>% 
  group_by(type) %>% 
  summarise(
    supper = sum(upper),
    slower = sum(lower),
    supper2 = sum(upper2),
    slower2 = sum(lower2),
    sum = sum(produced)
  )

uncertainty_crops_stats <- transform(uncertainty_crops_stats, row_stdev=apply(uncertainty_crops_stats, 1, sd, na.rm=TRUE))

uncertainty_crops_stats$Mean <- rowMeans(uncertainty_crops_stats[,2:5])

uncertainty_crops_stats

write_csv(uncertainty_crops_stats, "2_data/results/uncertainty_crops_stats.csv")
```


# 3) Other data weighing etc.

Total production, consumption and sales overview tables.

```{r}
produce_long <- readRDS("2_data/final/produce_long.RDS") %>% 
  weigh() %>% 
  mutate(across(c(produced:uncertain), ~. * weight_adj, .names = "w_{.col}")) 

write_csv(produce_long, "2_data/results/produce_long.csv")
# add households with no production for each type (TBD, could be another level... but this is the sankey level)
all_types <- produce_long %>% 
  select(category:item) %>% 
  unique()

allhhs <- weighting %>% 
  select(y4_hhid) 
```

Sales and destinations.

```{r}
b1_w <- readRDS("2_data/final/b1.RDS") %>% 
  weigh() %>% 
  mutate(w_sold = sold * weight_adj) 
```

Preharvest losses

```{r}
prelost <- readRDS("2_data/processed/crops_prelost.RDS") %>% 
  clear.labels() %>% 
  weigh() %>% 
  setDT()

write_csv(prelost, "2_data/results/prelost.csv")
```


```{r}
yieldgaps <- readRDS("2_data/processed/yieldgaps.RDS")
yieldgaps <- yieldgaps %>% right_join(weighting, by = "y4_hhid") %>% 
  mutate(
    pyg = yield*100/YP, # yield as percentage of potential
    w_area = area_planted_new * weight_adj, # total area
    w_harvest = yield * w_area, # total harvest expected for total area, in tonnes!
    w_yp = YP * w_area, # total yield potential for area
    w_yg = w_yp-w_harvest, # total yield gap for area in tonnes
    w_p_yield = w_harvest*100/w_yp # just for reference and checking
  )

write_csv(yieldgaps, "2_data/results/yieldgaps.csv")
```

# 4) MFA function

Add empty households

```{r}

```


Sankey function

```{r}
sankey <- function(data){
  nodes <- data.frame(name=c(as.character(data$source), as.character(data$target)) %>% unique())
  
  data$IDsource <- match(data$source, nodes$name)-1 
  data$IDtarget <- match(data$target, nodes$name)-1
  
  data <- data %>% 
    mutate_if(is.character, str_to_sentence)
  
  nb.cols <- nrow(nodes)
  mycolors <- colorRampPalette(brewer.pal(11, "PiYG"))(nb.cols)
  
  
  fig <- plot_ly(
    type = "sankey",
    orientation = "h",
    
    node = list(
      label = nodes$name,
      color = mycolors,
      pad = 15,
      thickness = 20,
      line = list(
        color = "black",
        width = 0.5
      )
    ),
    
    link = list(
      source = data$IDsource,
      target = data$IDtarget,
      value =  data$value,
      color = "light grey"
      
    )
  )
  # fig <- fig %>% layout(
  #   title = "Material Flow LSMS-ISA TZA wave 4",
  #   font = list(
  #     size = 10
  #   )
  #)
  fig
}
```

```{r}
# applicable to all
calc <- function(df) df[, lapply(.SD, sm), .SDcols=is.numeric, by=.(type)] # change function and grouping as required; for 3c this could stay item; use groupsgroups for other aggregation levels
# col sum
cl <- function(df) df[, lapply(.SD, sm), .SDcols=is.numeric] # change function as required
```

Keep in separate chunks and not as function until all certain and final.

## Crops
```{r}
cmfa <- copy(mfa_crops)

df <- cmfa[!is.na(type)] # households that were "added" for completion, they will probably need a type and item across all df with 0 values to represent the full sample in all dfs - TBD


  c1 <- calc(df)
  c2 <- cl(df)
  
  # first flow: harvest
  first <- c1[,. (source = type, value = produced)]
  first[, target := "Produced"]
  
  # second flow: harvest onward
  second <- c1[,.(source = type, 
                  consumed, 
                  sold, 
                  payment,
                  gifts,
                  losses, 
                  stored, 
                  feed, 
                  processing,
                  unallocated, 
                  missing)]
  
  second <- melt(second, id.vars = "source",
                 measure.vars = c("consumed", "sold", "payment", "gifts", "losses", "stored", "feed", "processing", "unallocated", "missing"))
  
  second <- upData(second,
                   rename = .q(variable = target))
  
  # third flow: processing
  third <- data.table(
    source = "processing",
    target = c("sold", "consumed", "waste"),
    value = c(pluck(c2$prodsold), pluck(c2$prodconsumed), pluck(c2$waste))
  )
  
  fourth <- data.table(
    source = "stored",
    target = "seed",
    value  = pluck(c2$seed)
  )

  
  data <- rbind(second, third, fourth)
  data <- data %>% select(target, source, value)

  sankey(data)
  
  
summary_crops <- c1 %>% 
  mutate(
    upper = produced + missing, # assumes production underestimated
    lower = produced - unallocated, # assumes production overestimated (no waste)
    mean = (produced + upper + lower)/3
  ) %>% 
  select(type, produced, upper, lower, mean, sold:uncertain, missing:unallocated)

cropsd <- summary_crops %>% select(upper, lower, produced)
cropsd <- transform(cropsd, row_stdev=apply(cropsd, 1, sd, na.rm=TRUE))

summary_crops$sd <- cropsd$row_stdev

summary_crops <- summary_crops %>% 
  relocate(sd, .after = mean)

write_csv(summary_crops, "2_data/results/summary_crops.csv")
```

## Animals

```{r}
df <- copy(mfa_animals)
df2 <- copy(mfa_hides)

# collapse and calculate
  m1 <- calc(df)
  m2 <- cl(df)
  m22 <- cl(df2)
  
  # feed flow to total slaughter weight
  feed <- melt(m1, id.vars = "type",
               measure.vars = c("feed", "grazed"))
  feed <- upData(feed, rename = .q(variable = source, type = target))
  
  # first flow types to slaughter
  first <- m1[,.(source = type, value = total_weight)]
  first[, target := "slaughter" ]
  
  # second flow from slaughter
  second <- data.table(
    source = "slaughter",
    target = c("sold", "meat", "offal", "hides", "inedible"), # could include uncertain
    value = c(pluck(m2$sold_weight), pluck(m2$meat), pluck(m2$offal), pluck(m2$hides), pluck(m2$inedible))
  )

  
  # last flow from hides
  # meat, offal to consumed
  # hides to unused and processed
  # processed hides to sold and consumed
  third <- data.table(
    source =  c("meat", "offal"),
    target = "consumed",
    value = c(pluck(m2$meat), m2$offal)
  )
  
  fourth <- data.table(
    source = c("hides","processing", "processing", "hides"),
    target = c("processing", "sold", "unallocated", "missing","unused"),
    value = c(pluck(m22$pprod), pluck(m22$sold2), pluck(m22$unallocated), pluck(m22$missing), pluck(m2$hides)-pluck(m22$hides))
  )
  
  
  data2 <- rbind(feed, first, second, third, fourth)
  
  sankey(data2)
  
summary_animals <- m1 %>% 
  mutate(
    upper = total_weight + missing, # assumes production underestimated
    lower = total_weight - unallocated, # assumes production overestimated (no waste)
    mean = (total_weight + upper + lower)/3
  ) %>% 
  select(type:grazed, produced = total_weight, upper, lower, mean, sold = sold_weight, consumed = cons_weight, meat, offal, hides, inedible, missing, unallocated)

animalsd <- summary_animals %>% select(upper, lower, produced)
animalsd <- transform(animalsd, row_stdev=apply(animalsd, 1, sd, na.rm=TRUE))

summary_animals$sd <- animalsd$row_stdev

summary_animals <- summary_animals %>% 
  relocate(sd, .after = mean)

write_csv(summary_animals, "2_data/results/summary_animals.csv")
```

## Milk
```{r}
df <- copy(mfa_milk)
# calculate
  mi1 <- calc(df)
  mi2 <- cl(df)
  
  # feed
  feed <- melt(mi1, id.vars = "type",
               measure.vars = c("feed", "grazed"))
  feed <- upData(feed, rename = .q(variable = source, type = target))
  
  # feed to production
  firsti <- data.table(
    source = c("large ruminants", "small ruminants"),
    target = "milked",
    value = c(pluck(mi1[1,4]), pluck(mi1[2,4])) # shaky as code..
  )

  
  # first flow production onward
  first <- mi1[,.(source = type, 
                  consumed, 
                  sold, 
                  processed, 
                  missing, 
                  unallocated)]
  
  first <- melt(first, id.vars = "source",
                measure.vars = c("consumed", "sold", "processed", "missing", "unallocated"))
  first <- upData(first,
                  rename = .q(variable = target),
                  source = "milked")
  
  # second flow processing to sold & consumed
  second <- data.table(
    source = 
      "processed",
    target = 
      c("sold", "unallocated"),
    value = c(
      pluck(mi2$psold), pluck(mi2$processed)-pluck(mi2$psold))
  )
  
  data3 <- rbind(feed, firsti, first, second)
  sankey(data3)
  
summary_milk <- mi1 %>% 
    mutate(
    upper = produced + missing, # assumes production underestimated ! differs from range milk calculations which uses lower and upper milk production
    lower = produced - unallocated, # assumes production overestimated (no waste)
    mean = (produced + upper + lower)/3
  ) %>% 
  select(type:grazed, produced, upper, lower, mean, consumed:psold, missing:unallocated)

milksd <- summary_milk %>% select(upper, lower, produced)
milksd <- transform(milksd, row_stdev=apply(milksd, 1, sd, na.rm=TRUE))

summary_milk$sd <- milksd$row_stdev

summary_milk <- summary_milk %>% 
  relocate(sd, .after = mean)

write_csv(summary_milk, "2_data/results/summary_milk.csv")
```

## Eggs
```{r}
df <- copy(mfa_eggs) %>% 
  filter(!is.na(item)) 

e1 <- calc(df)
  
data4 <- data.table(
    source = c(
      "feed","grazed", 
      "poultry",
      "eggs", "eggs", "eggs"),
    target = c(
      "poultry", "poultry",
      "eggs",
      "sold", "consumed", "unallocated"),
    value = c(
      pluck(e1$feed), pluck(e1$grazed), 
      pluck(e1$produced),
      pluck(e1$sold), pluck(e1$consumed), pluck(e1$unallocated))
  )

  
  sankey(data4)
  
summary_eggs <- e1 %>% 
  mutate(
    lower = produced - unallocated, # assumes production overestimated (no waste)
    mean = (produced + lower)/2
  ) %>% 
  select(type:grazed, produced, lower, mean, consumed, missing:unallocated)

eggssd <- summary_eggs %>% select(lower, produced)
eggssd <- transform(eggssd, row_stdev=apply(eggssd, 1, sd, na.rm=TRUE))

summary_eggs$sd <- eggssd$row_stdev

summary_eggs <- summary_eggs %>% 
  relocate(sd, .after = mean)

write_csv(summary_eggs, "2_data/results/summary_eggs.csv")
```

## All
```{r}
mfa <- rbind(data, data2, data3, data4)

upcase <- function(df){
  df <- df %>% 
    mutate_if(is.character, str_to_sentence) %>% 
    mutate_if(is.factor, function(x) str_to_sentence(as.character(x)))
}

mfaw <- mfa %>% 
  upcase() %>% 
  mutate(target = ifelse(target == "Consumed", "Household", target),
         source = ifelse(source == "Consumed", "Household", source),
         target = ifelse(target == "Sold", "Sales", target),
         source = ifelse(source == "Sold", "Sales", source),
         target = ifelse(target == "Processing", "Processed", target),
         source = ifelse(source == "Processing", "Processed", source),
         # new groupings - combine targets
         target = ifelse(
           target == "Waste" | target == "Losses" | target == "Inedible" | target == "Unused", "Loss & Waste", target),
         # target = ifelse(
         #   target == "Missing", "Uncertain", target
         # ),
         value = as.numeric(value)) %>% 
  group_by(source, target) 

write_csv(mfaw, "2_data/final/mfaw.csv")
write_csv(mfaw, "2_data/results/mfaw.csv")

sankey(mfaw)
```

# Input for STAN

```{r eval=FALSE}
# generate new uncertainty estimates that are weighted for crops and for milk
uncertainty_crops_weighted <- crops %>% 
  weigh() %>% 
  mutate(
    uncertain = uncertain * weight_adj,
    produced = produced * weight_adj,
    upper = ifelse(uncertain<0, uncertain*-1+produced, produced), # production insufficient to cover all disposition
    lower = ifelse(uncertain>0, produced-uncertain, produced),# production exceeds disposition, we assume production is lower than reported (this can never be true as the list of uses asked about is limited)
    upper2 = ifelse(uncertain2<0, uncertain2*-1+produced, produced), # same as above, assume processing is a separate entity
    lower2 = ifelse(uncertain2>0, produced-uncertain2, produced),
    meanval = (upper+lower+upper2+lower2+produced)/5) %>% 
  group_by(type) %>% 
  summarise(
    supper = sum(upper),
    slower = sum(lower),
    supper2 = sum(upper2),
    slower2 = sum(lower2),
    sum = sum(produced)
  ) 

uncertainty_crops_weighted <- transform(uncertainty_crops_weighted, row_stdev=apply(uncertainty_crops_weighted, 1, sd, na.rm=TRUE))

uncertainty_crops_weighted$Mean <- rowMeans(uncertainty_crops_weighted[,2:5])

uncertainty_crops_weighted <- uncertainty_crops_weighted %>% 
  mutate(mean = Mean/t,
         sd = row_stdev/t) %>% 
  select(type, mean, sd) %>% 
  rd()

uncertainty_milk_weighted <- milk %>% 
  weigh() %>% 
  mutate(
    uncertain = uncertain*weight_adj,
    produced = produced*weight_adj,
    upper = ifelse(uncertain<0, uncertain*-1+produced, produced), # production insufficient to cover all disposition
    lower = ifelse(uncertain>0, produced-uncertain, produced) # production lower than all disposition
  ) %>% 
  group_by(type) %>% 
  summarise(
    supper = sum(milk_hi),
    slower = sum(milk_lo),
    supper2 = sum(upper),
    slower2 = sum(lower),
    sum = sum(produced)
  ) 

uncertainty_milk_weighted <- transform(uncertainty_milk_weighted, row_stdev=apply(uncertainty_milk_weighted, 1, sd, na.rm=TRUE))

uncertainty_milk_weighted$Mean <- rowMeans(uncertainty_milk_weighted[,2:5])

uncertainty_milk_weighted <- uncertainty_milk_weighted %>% 
  mutate(mean = Mean/t,
         sd = row_stdev/t) %>% 
  select(type, mean, sd) %>% 
  rd()

# use remaining from mfaw file as below
```


```{r}
uncertain_joined <- uncertainty_crops_stats %>% 
  select(type, mean = Mean, sd = row_stdev) %>% 
  mutate(
    mean = mean/t,
    sd = sd/t
  ) %>% 
  rd()

crops_simple <- mfaw %>% 
  filter(grepl("Grains|Legu|Nuts|Roots|Fruits|Veg|crops|Spices", source)) %>% 
  group_by(target) %>% 
  summarise(value = sm(value)/t) %>% 
  add_column(source = "Harvest")

processed2 <- mfaw %>% 
  filter(target == "Processed" & (source== "Milked" | source=="Hides")) %>% 
  group_by(target, source) %>% 
  summarise(value = sm(value)/t) 

processedout <- mfaw %>% 
  filter(source == "Processed") %>% 
  group_by(source, target) %>% 
  summarise(value = sm(value)/t) 

graze <- mfaw %>% 
  filter(source == "Grazed") %>% 
  group_by(target) %>% 
  summarise(value = sm(value)/t) %>% 
  add_column(source = "Grazed")

feed <- mfaw %>% 
  filter(source == "Feed") %>% 
  group_by(target) %>% 
  summarise(value = sm(value)/t) %>% 
  add_column(source = "Feed")

slaughter <- mfaw %>% 
  filter(target == "Slaughter") %>% 
  group_by(source) %>% 
  summarise(value = sm(value)/t) %>% 
  add_column(target = "Slaughter")

slaughterout <- mfaw %>% 
  filter(source == "Slaughter") %>% 
  group_by(target) %>% 
  summarise(value = sm(value)/t) %>% 
  add_column(source = "Slaughter")

milkout <- mfaw %>% 
  filter(source == "Milked") %>% 
  group_by(target) %>% 
  summarise(value = sm(value)/t) %>% 
  add_column(source = "Milked")

grazedin <- mfaw %>% # fictional
  filter(source == "Grazed") %>% 
  group_by(source) %>% 
  summarise(value = sm(value)/t)
```


