---
title: "03_Animal_products"
author: "Heike Rolker"
date: "2024-01-10"
output: html_document
---

```{r setup, include=FALSE}
library(here)
source(here::here("01-smallholder-material-flow", "scripts", "packages.R"))
source(here::here("01-smallholder-material-flow", "scripts", "functions.R"))
```

# Other products

Overview:

Animal products, except honey.

- **produce** all animal products, cleaned file but mostly raw data

```{r produce hk}
lf_sec_08 <- read_dta(here::here("data", "raw", "lsms","lf_sec_08.dta"))

lapply(lf_sec_08, attr, "label")
```

```{r eggs naming}
produce <- clean_up(lf_sec_08)
produce <- upData(produce,
               rename = .q(
                 lf08_02 = length,
                 lf08_03_1 = produced,
                 lf08_03_2 = unit,
                 lf08_05_1 = sold,
                 lf08_05_2 = unitsold,
                 lf08_06 = value,
                 lf08_07_1 = buyer1,
                 lf08_07_2 = buyer2
               ),
               # true 0 when NA
               produced = ifelse(lf08_01 == "no", 0, produced),
               sold = ifelse(lf08_04 == "no", 0, sold),
               sold = as.numeric(sold),
               
               labels = .q(
                 length = "Number of months produced in past 12 mo",
                 produced = "Average quantity produced per month",
                 unit = "Unit of production",
                 sold = "Quantity sold in last 12 months",
                 value = "Total value of sales in past 12 months"
               ))

saveRDS(produce, file = here::here("data", "processed", "01", "produce.RDS"), compress = TRUE)
```

### Hides

Overview:

- **hides** overview of hides data matching to rums where possible based on ownership, does not account for quantities allocated, contains buyers

- **mass_hides_long** extended information on hides at household level, including those not from slaughter, duplication as some not matched to animal types

- **mass_hides** only hide from ruminants and complete information

Only hides from slaughtered animals are relevant as others outside of the system boundary.

Aim:

-   prepare data for matching

-   all into kg

```{r prep slaughter}
produce <- readRDS(here::here("data", "processed", "01", "produce.RDS"))
# extract hides
hides <- produce[productid == "skin / hides"]
# slaughter info
wa <- readRDS(here::here("data", "processed", "01", "wa.RDS")) # just not to alter too much code...

# determine hides origin, match with slaughter info
ah <- wa[,.(y4_hhid, type, lvstckid, slaughter, hides, ssold)]
ah[, slaughter := slaughter - ssold] # only animals slaughtered for assumed consumption, i.e. those that have a hides estimate
ah[, slaughter := adlab(slaughter, "Animals slaughtered for assumed consumption only")]

# collapse groups to hh level to match produce recording
ah <- ah[, lapply(.SD, sm), .SDcols=is.numeric, by=.(y4_hhid, type)]

# flip data table (a lot of work around)
sl <- ah %>% 
  spread(type, slaughter) %>% 
  select(!c(poultry, 'other animals', pigs, hides)) %>% 
  dplyr::rename(lgrm = `large ruminants`) %>% 
  dplyr::rename(smrm = `small ruminants`) %>% 
  mutate(
    smrm = as.numeric(replace_na(smrm,0)),
    lgrm = as.numeric(replace_na(lgrm, 0))
  ) %>% 
  group_by(y4_hhid) %>% 
  summarise(
    smrm = sum(smrm),
    lgrm = sum(lgrm)
  )

hi <- ah %>% 
  select(!slaughter) %>% 
  spread(type, hides) %>% 
  select(!c(poultry, 'other animals', pigs)) %>% 
  dplyr::rename(hlgrm = `large ruminants`) %>% 
  dplyr::rename(hsmrm = `small ruminants`)

# merge hides with slaughter and hide weight data
hides <- merge(hides, sl, by='y4_hhid', all=TRUE) 
hides <- merge(hides, hi, by='y4_hhid', all=TRUE) 

# annual producition of pieces
hides[, `:=` (pieces = ifelse(unit == "pieces", produced*length, NA),
              total = ifelse(unit == "kgs" | unit == "litres", produced * length, NA))] 

# align type (should be easier and in command above..)
hides[, `:=` (pieces = as.double(pieces),
              total = as.double(total))]

hides[, `:=` (weight = fcase(
                   pieces <= lgrm, hlgrm, # large rums
                   pieces <= smrm, hsmrm, # small rums
                   pieces > smrm & (is.na(lgrm) | lgrm == 0), hsmrm/pieces*smrm, # small rums
                   pieces > lgrm & (is.na(smrm) | smrm == 0), hlgrm/pieces*lgrm, # large rums
                   pieces >= (smrm + lgrm), hsmrm + hlgrm, # only hides from slaughter are counted
                   total <= hlgrm & hsmrm == 0, total,      # all hides allocated to large rums
                   total <= hsmrm & hlgrm == 0, total,      # all hides allocated to small rums
                   total >  hlgrm & hsmrm == 0, hlgrm,      # all hides allocated to large rums
                   total >  hsmrm & hlgrm == 0, hsmrm,       # all hides allocated to small rums
                   total < (hsmrm + hlgrm), total,          # hides not allocated but sufficient
                   total > (hsmrm + hlgrm), hsmrm + hlgrm), # only hides from slaughter are counted
              # allocate the hides to animal types where possible
              # where type is NA, either all hides from both types are used or a fraction of both
              type = fcase(
                pieces <= lgrm, "large ruminants",
                pieces <= smrm, "small ruminants",
                pieces > smrm & (is.na(lgrm) | lgrm == 0), "small ruminants",
                pieces > lgrm & (is.na(smrm) | smrm == 0), "large ruminants",
                total <= hlgrm & hsmrm == 0, "large ruminants",
                total <= hsmrm & hlgrm == 0, "small ruminants",
                total >= hlgrm & hsmrm == 0, "large ruminants",
                total >= hsmrm & hlgrm == 0, "small ruminants"))]

# sales of hides
hides[, perc_sold := fcase(
                unit == "kgs" | unit == "litres", sold/total, # not all use the same units but looks like error
                unit == "pieces", sold/pieces)]

# weight of sold hides
hides[, weight_sold := ifelse(sold == 0, 0, perc_sold * weight)]
# replace NA with 0
hides[, weight_sold := ifelse(is.na(weight_sold), 0, weight_sold)]

# determine hides not sold assumed consumed
hides[, missing := weight - weight_sold]

# change labels
hides <- upData(hides,
              labels = .q(
                total = "Total quantity produced annually derived from kg or litre", # let's pretend it's the same
                perc_sold = "Percentage of skins/hides production sold",
                pieces = "Annual production of hides in pieces as reported",
                consumed = "Skins/hides produced but not sold",
                missing = "Quantities unaccounted, assumed consumed"
              ), 
              units = .q(
                total = kg,
                weight_sold = kg
              ), 
              drop = .q(
                smrm, # number of small rum slaughtered
                lgrm, # number of large rum slaughtered
                hsmrm, # weight of hides from slaughtered small rums
                hlgrm, # weight of hides from slaughtered large rums
                lf08_08_2, # not needed
                lf08_08_1,
                lf08_06, # not needed
                unit, # converted to kg
                unitsold, # converted to kg
                lf08_04, # sales yes/no
                length, # total computed
                produced, # total computed
                lf08_01, # produced yes/no
                sold # not needed
              )
              )

# split unallocated types between small and large rums (BONUS if strictly necessary)

hides <- upData(hides,
                rename = .q(
                  weight = produced,
                  weight_sold = sold
                ),
                drop = .q(
                  pieces,
                  total,
                  perc_sold
                ))

hides <- hides[produced>0]
# remaining & missing conditions
# NOTE: diagnostic checks — retained for reference, not part of pipeline output
# hid[produced > 0 &is.na(type)]
# hid[unitsold != "pieces" & total < sold]
# hid[unitsold == "pieces" & pieces < sold]
# estimate weight of hides sold or used for own consumption derived from slaughtered animals

saveRDS(hides, here::here("data", "processed", "01", "hides.RDS"), compress = TRUE)
```

*All weight of hides that can be allocated to slaughtered animals are allocated to them, the remainder is ignored. Method not ideal as it standardizes the quantities of hide per animal. Hides from both types not possible.*

Aim (BONUS, *a lot of effort for marginal gains, probably sufficient to allocate them to animal type if anything*):

-   allocate the hides produced to the animals slaughtered

manualfix\[sums, on = (y4_hhid)\]

REVIEW MERGES IN CHUNK BELOW, CREATES DUPLICATION FOR ENTRIES WITH EXCLUSION → solved

Not sure what to do with code below... should just allocate hides to animal types not contain as much information as it presently does...


```{r slaughter hides}
# reshape wa for hides, edible, inedible to become
slaughter <- wa[slaughter > 0]

# ASSUMPTION: HIDES ARE ONLY PRODUCED FROM ANIMALS SLAUGHTERED AND NOT SOLD (these are the only hides of interest)
# select info for animals slaughtered
meat <- slaughter[,.(y4_hhid, type, lvstckid, slaughter, weight, hides, sold_weight)] # sold_weight needed?

# select only full info from hides
hime <- hides[,.(y4_hhid, type, pprod = produced, sold2 = sold, missing)]
hime <- hime[pprod > 0]

# collapse meat to type level to match level from hides
meatt <- meat[, lapply(.SD, sm), .SDcols=is.numeric, by=.(y4_hhid, type)]

# merge slaughter info (meat) to hides info (hime) at animal type and hh level
himeatt <- merge(hime, meatt, by = c("y4_hhid", "type"), all = TRUE)

# extract hhs with no type to add animal type manually
hides_hh <- himeatt[is.na(type)]

# remove empty vars because no match in type (is missing)
hides_hh <- hides_hh[,.(y4_hhid, pprod, sold2, missing)]

# extract rums only for matching manually to hides_hh
rumsonly <- meatt[type == "large ruminants" | type == "small ruminants"]

# rematch with rums only (will result in duplication)
manualfix <- rumsonly[hides_hh, on=.(y4_hhid)]

# split hides produced across animal types
sums <- manualfix[, lapply(.SD, sm), .SDcols=is.numeric, by=.(y4_hhid)]
sums[,.(y4_hhid, hides)]
hides_hh[sums, on = .(y4_hhid)] # check how many have produced more/less hides than sum of hides by type

manualfix[, `:=` (pprod_new = hides,
                  perc_sold = sold2/pprod/2)]

# three households for which above is not applicable and require manual recoding
# Manual fix: three households where hides cannot be split between large/small ruminants
# algorithmically — values assigned by inspection of raw data.
# 7294-001: equal split across both types (3 animals total → 1.5 each)
# 8014-001: all 5 hides from small ruminants, 0 from large
# 4764-001: equal split (2 animals total → 1 each)
# Flag for review if raw data or slaughter records change.
manualfix$pprod_new[which(manualfix$y4_hhid == "7294-001" & manualfix$type == "large ruminants")] <- 1.5
manualfix$pprod_new[which(manualfix$y4_hhid == "7294-001" & manualfix$type == "small ruminants")] <- 1.5

manualfix$pprod_new[which(manualfix$y4_hhid == "8014-001" & manualfix$type == "large ruminants")] <- 0
manualfix$pprod_new[which(manualfix$y4_hhid == "8014-001" & manualfix$type == "small ruminants")] <- 5

manualfix$pprod_new[which(manualfix$y4_hhid == "4764-001" & manualfix$type == "large ruminants")] <- 1
manualfix$pprod_new[which(manualfix$y4_hhid == "4764-001" & manualfix$type == "small ruminants")] <- 1

manualfix[, sold_new := perc_sold* pprod_new]

# rename and delete 
manualfix[,.q(pprod, sold2, perc_sold) := NULL]

manualfix <- upData(manualfix,
                    rename = .q(
                      sold_new = sold2,
                      pprod_new = pprod
                    ))

# replace na with 0
manualfix[, sold2 := replace_na(sold2, 0)]

# recalculate missing
manualfix[, missing := pprod-sold2] # hides not sold assumed for household


### Aim:putting it all back together


# anti join manualfix with himeatt to avoid duplication
new <- dplyr::anti_join(himeatt, manualfix, by = c("y4_hhid", "type"))
# NOTE (backlog): new2 is a data.table equivalent of new — same result, not used downstream.
# Remove once pipeline is stable and standardise to data.table anti-join throughout.
new2 <- himeatt[!manualfix, on =.(y4_hhid, type)]

# match order to overcome binding error
new <- new[,.(y4_hhid, type, weight, hides, sold_weight, pprod, sold2, missing)]
new[, missing := as.double(missing)]

# bind rows manualfix and himeatt
meat <- bind_rows(manualfix, new)

# add feed and exclusion information to meat 
meat_sl <- slaughter[,.(type, y4_hhid, meat, offal, hides, inedible)]
# ex[, ex := ifelse(is.na(excl),0,1)] # same household excluded multiple times? --> no
# reduce to type level of hides
meat_sl <- meat_sl[, lapply(.SD, sm), .SDcols=is.numeric, by=.(y4_hhid, type)]

# add exclusions back in -> exclusions are seperate.
# NOTE: diagnostic checks — retained for reference, not part of pipeline output
# ex <- af[,.(type, y4_hhid, excl)]
# ex <- ex %>% filter(!is.na(excl)) %>% unique()
# ex <- ex[meat_sl, on = .(type,  y4_hhid)]

# merge all slaughter/hides info
# meat2 <- ex[meat, on = .(type,  y4_hhid)] # introduces duplication
# meat <- meat2[!is.na(type)] # a number unmatched -> ignore!

# label all vars
mass_hides_long <- upData(meat,
               labels = .q(
                 weight = "Total weight of animals slaughtered",
                 edible = "Estimate of edible product",
                 inedible = "Estimate of inedible product",
                 hides = "Estimate of total hides",
                 sold_weight = "Live weight of animals slaughtered",
                 # feed = "Feed required for meat production",
                 # grazed = "Roughage required for meat production",
                 pprod  = "Weight of hides produced from slaughter",
                 sold2 = "Weight of hides sold",
                 missing = "Hides not sold (assumed for household)"))

mass_hides <- mass_hides_long %>% 
  left_join(select(hides, y4_hhid, produced), by = c("y4_hhid")) %>% 
  mutate(
    rel_prod = pprod/produced,
    rel_prod = ifelse(produced>0 & is.na(rel_prod), 0, rel_prod)
  ) %>% 
  select(!c(produced, weight, sold_weight)) %>% 
  filter((type == "small ruminants" | type == "large ruminants") & !is.na(pprod))

mass_hides[, rel_prod := adlab(rel_prod, "Proportion of hides produced from total (i.e. not slaughter)")]

saveRDS(mass_hides, here::here("data", "processed", "01", "mass_hides.RDS"), compress = T)
saveRDS(mass_hides_long, here::here("data", "processed", "01", "mass_hides_long.RDS"), compress = T)
```

To merge hides with meat there are two options:

-   collapse meat to the type level to merge with hides (all hide source is an estimate anyhow), and split the few entries between large and small rums manually

-   distribute type from hides across the animals slaughtered in meat (more cumbersome) --\> no


### Eggs

Final files:

**excl_eggs** list of exclusions for eggs

**mass_eggs** production and destination estimates for eggs

```{r eggs animals}
animals_fin <- readRDS(here::here("data", "processed", "01", "animals_fin.RDS"))
f <- readRDS(here::here("data", "processed", "01", "feed_short.RDS"))

chicken <- data.table( # chicken p. 107 backyard
  feed1 = c("only feeding (no grazing/scavenging)", 
            "mainly grazing/scavenging w/ some feeding", 
            "only grazing/scavenging", 
            "mainly feeding w/ some grazing/scavenging", 
            "tethering"),
  type = "poultry",
  feed = c(1, 0.4, 0, 0.6, 0.5), # cross-refed with categorisation in FAO
  grazed = c(0, 0.6, 1, 0.4, 0.5))

# check plausible for poultry only
ae <- animals_fin[,.(y4_hhid, type, lvstckid, current, max_owned)]
ae <- ae[type=="poultry"]

# ae[lvstckid == "other poultry" & current > 0]
# ae[lvstckid == "other poultry" & max_owned > 0]
# ae[lvstckid == "other poultry"]

# collapse to hh level
ae <- ae[, lapply(.SD, sm), .SDcols=is.numeric, by=.(y4_hhid, type)]
```

```{r feeds table, eval=FALSE}
# NOTE: eval=FALSE — feedtbl combines milkfeed and chicken feed tables.
# milkfeed is not defined in this script — requires milk wrangling to run first.
# Re-enable once milk section is integrated into pipeline.
milkfeed <- milkfeed %>% 
  mutate(
    type = ifelse(type == "small ruminants", "milk (small ruminants)", "milk (large ruminants)"))

chicken <- chicken %>% 
  mutate(type = "eggs")

feeds

feeds <- clear.labels(feeds)

feedtbl <- bind_rows(milkfeed, chicken, feeds)


# purpose? for appendix?
saveRDS(feedtbl, here::here("data", "processed", "01", "feedtbl.RDS"), compress = T)
```

Aim:

-   determine total weight of eggs produced

```{r eggs}
produce <- readRDS(here::here("data", "processed", "01", "produce.RDS"))

# adjust class
produce[, produced := as.numeric(produced)]
produce[, sold := as.numeric(sold)]

# extract eggs
e <- produce[productid == "eggs"]

# convert all to kg
e <- upData(e,
       produced_new = fcase(
         unit == "kgs", produced * length,
         unit == "pieces", produced * 41.26/1000 * length, # egg weight 41.26g for SSA based on [@MacLeod.2013] p.105
         unit == "litres", produced *1.03 * length
        # https://www.click-chef.com/?lang=3: 1l = 940g
        # http://convert-to.com/720/chicken-eggs-5-sizes-conversion-plus-nutrients-values.html: 1l = 1.03 kg -> includes shell
       ),
       sold_new =fcase( # asked across 12 months
          unitsold == "kgs", sold,
          unitsold == "pieces", sold * 41.26/1000, 
          unitsold == "litres", sold *1.03),
       sold_new = ifelse(sold == 0, 0, sold_new),
       n_eggs = produced_new/0.04126/length,
       N_eggs = produced_new/0.04126,
       labels = .q(
         produced_new = "Annual quantity produced",
         sold_new = "Annual quantity sold",
         consumed = "Quantity consumed",
         n_eggs = "Number of eggs per month",
         N_eggs = "Total quantity of eggs produced"
       ),
       units = .q(
         produced_new = kg,
         sold_new = kg
       ))

# Replace NA with 0 in computed egg quantity columns — named to avoid positional fragility
cols_to_zero <- c("produced_new", "sold_new", "n_eggs", "N_eggs")
e[, (cols_to_zero) := lapply(.SD, function(x) ifelse(is.na(x), 0, x)), .SDcols = cols_to_zero]


```

Aim (BONUS):

-   determine destinations

Aim:

-   exclude implausible values

```{r eggs exclusions}
eggs <- e[,.(y4_hhid, productid, produced = produced_new, sold = sold_new, n_eggs, N_eggs, length)]

# match with animals owned
eggs <- ae[eggs, on = .(y4_hhid)]

# top 2% of eggs per month
top <- top2(eggs$n_eggs) 

# top 2% eggs produced across mean
eggs$mean <- rowMeans(eggs[,c('current', 'max_owned')], na.rm=TRUE) # mean of currently owned and max owned, assumes true value somewhere in between
eggs[, n_ea := n_eggs/mean] # eggs/month per poultry based on max

eggs[, n_ea := adlab(n_ea, 'Eggs per month per estimated n of poultry owned')]
topa <- top2(eggs$n_ea) # top 2% of eggs produced per month per estimated number of animals owned

# estimate number of poultry needed per month
eggs[, n_est := n_eggs/45]
eggs[, n_est := adlab(n_est, 'Est number of poultry required')]

excl_eggs <- eggs[, excl := fcase(
    # top<n_eggs, "Top 2%", # difficult to justify if i do not know how many animals lay eggs at any given time
    sold>produced, "Implausible",
    n_est>mean, "Insufficient animals" # estimated number of animals exceeds estimated ownership
    )]

excl_eggs <- excl_eggs %>% 
  select(y4_hhid, item = productid, excl)
# estimate number of eggs per animal
# Number of eggs/hen/year : 45; @MacLeod.2013 p.105

# eggs[n_est>max_owned]
# eggs[n_est>current]
# eggs[n_est>mean]
write.csv(excl_eggs, here::here("data", "processed", "01", "excl_eggs.csv"))
```

Aim:

-   determine feed requirements

```{r eggs feed}
ef <- eggs[,.(y4_hhid, type, productid, produced, sold, N_eggs)]

# merge with feed
ef <- f[ef, on = c("type", "y4_hhid")]

# merge with feed type used (same as above)
ef <- chicken[ef, on = c("feed1", "type")]

mass_eggs <- upData(ef,
             need = produced * 2.3, # based on @Alexander.2016
             feed = need * feed,
             grazed = need * grazed,
             missing = produced-sold,
             consumed = 0, # place holder
             labels = .q(
               need = "Feed requirements in DM",
               feed = "Quantity of feed consumed in DM",
               grazed = "Quantity grazed/scavenged in DM"
             ),
             drop = .q(
               feed1, need, N_eggs
             )
  
)

saveRDS(mass_eggs, here::here("data", "processed", "01", "mass_eggs.RDS"), compress = TRUE)
```


### Estimate egg consumption

Below requires a consistent approach that is defensible.

Aim:

-   estimate quantity of eggs consumed based on recall info

Egg consumption statistics from household recall data.

```{r}
# Allocate household egg production between consumption and unallocated flow using
# combined agricultural and household survey data.
# ---- 1. Load Data ----
# Egg production and sales: from agricultural survey (annualized at this stage)
prod <- readRDS(here::here("data", "processed", "01", "mass_eggs.RDS")) # y4_hhid, produced, sold, missing (i.e., unallocated)
# Household 7-day recall, all sources
recall <- readRDS(here::here("data", "processed", "02", "recall_kg.RDS")) %>% clear.labels() 
# Ag households only
ag_ids <- readRDS(here::here("data", "processed", "01", "ag_ids.RDS")) %>% clear.labels()
# Household size from LSMS
hhsize <- haven::read_dta(here::here("data", "raw", "lsms", "consumptionNPS4.dta")) %>%
  dplyr::select(y4_hhid, hhsize) %>%
  left_join(select(ag_ids, y4_hhid), by = "y4_hhid")

# ---- 2. Compute recall-based weekly egg consumption ----
recall_eggs <- recall %>%
  filter(itemcode == "eggs") %>%
  left_join(hhsize, by = "y4_hhid") %>%
  mutate(
    quantity_kg = ifelse(consumed == "no", 0, quantity_kg),
    hhsize = ifelse(is.na(hhsize), NA_real_, hhsize)
  ) %>%
  group_by(y4_hhid) %>%
  summarise(
    recall_weekly_kg = sum(quantity_kg, na.rm = TRUE),   # total household weekly consumption
    hhsize = first(hhsize)
  ) %>%
  ungroup() 

# ---- 3. Calculate national mean per-person weekly egg consumption (AG HHs Only) ----
national_mean_ppwk <- recall_eggs %>%
  filter(recall_weekly_kg > 0, !is.na(hhsize)) %>%
  summarise(mean_ppwk = mean(recall_weekly_kg / hhsize, na.rm = TRUE)) %>%
  pull(mean_ppwk)

# ---- 4. Merge everything into allocation table ----
eggs_alloc <- prod %>%           # y4_hhid, produced, sold, missing (i.e. unallocated)
  left_join(recall_eggs, by = "y4_hhid") %>%
  mutate(
    hhsize = haven::as_factor(hhsize),     # if it’s a labelled vector
    hhsize = as.numeric(as.character(hhsize)),
    weeks_produced = (produced + sold + missing) / 7, # Approximate number of weeks with egg flows; adjust as needed
    annualized_recall_kg = recall_weekly_kg * 52,     # Extrapolate recall (if present) to full year
    annualized_impute_kg = national_mean_ppwk * hhsize * 52 # fallback: national mean * hhsize * weeks/year
  )

# ---- 5. Encapsulate rolling allocation logic ----
allocate_consumed <- function(annualized_recall_kg, missing, annualized_impute_kg) {
  # Returns (consumed, method)
  if (!is.na(annualized_recall_kg) && annualized_recall_kg <= missing) {
    return(list(consumed = annualized_recall_kg, method = "recall-extrapolated"))
  } else if (!is.na(annualized_impute_kg) && annualized_impute_kg <= missing) {
    return(list(consumed = annualized_impute_kg, method = "mean-imputed"))
  } else if (!is.na(missing) && missing > 0) {
    return(list(consumed = missing, method = "capped-at-unallocated"))
  } else {
    return(list(consumed = NA_real_, method = "not-allocated"))
  }
}

tmp <- purrr::pmap_dfr(
  eggs_alloc[,c("annualized_recall_kg","missing","annualized_impute_kg")],
  allocate_consumed
) %>% dplyr::rename(
  consumed_alloc = consumed,
  allocation_method = method
)

eggs_alloc_final <- eggs_alloc %>%
  bind_cols(tmp) %>%
  mutate(
    unallocated = missing - consumed_alloc,
    unallocated = ifelse(unallocated < 0, 0, unallocated)
  ) %>%
  select(
    y4_hhid, produced, sold, missing,
    consumed = consumed_alloc, allocation_method, unallocated, recall_weekly_kg, hhsize
  )

# ---- 6. QA Checks ----
n_fails <- sum(is.na(eggs_alloc_final$consumed))
if (n_fails > 0) {
  message(n_fails, " households could not be allocated consumption using recall or imputation.")
  write.csv(eggs_alloc_final %>% filter(is.na(consumed)),
            here::here("data", "processed", "01", "egg_allocation_failures.csv"),
            row.names = FALSE)
}

# ---- 7. Save Output ----
saveRDS(eggs_alloc_final, here::here("data", "processed", "01", "mass_eggs_cons.RDS"), compress = TRUE)

# For Tableau/Excel QA or exploration
readr::write_csv(eggs_alloc_final, here::here("data", "processed", "01", "mass_eggs_cons.csv"))
```

Overview for computation of egg consumption

1) based on weekly recall, egg consumption is extrapolated to annual consumption (regardless of source), if extrapolated consumption is less than missing when sales are subtracted from production, the estimated annual consumption (from weekly) is assumed to have been consumed (the remainder stays missing)

2) if household didn't consume in recall week but has missing quantities, we use the mean per person consumption across households, extrapolate it to annual consumption and multiply by number of household members, if this assumed annual consumption is less than the missing quantities, the estimate is assumed to have been consumed and the remainder stays missing

3) the household did consume in the recall week, but missing is smaller than what we would assume to have been consumed, thus we allocate all missing to consumption (resulting in no missing for these households, around n = 144) -> this is a rather weak assumption as there MUST be missing/loss




