---
title: "04_Milk"
author: "Heike Rolker"
date: "2024-02-09"
output: html_document
---

```{r setup, include=FALSE}
library(here)
source(here::here("01-smallholder-material-flow", "scripts", "packages.R"))
source(here::here("01-smallholder-material-flow", "scripts", "functions.R"))
```

Livestock survey - milk & all animal products

Final files:

- **milk** overview milk, cleaned data in metric

- **excl_milk** exclusions for milk

- **mass_milk** final masses for milk, destinations and feed requirements


# Milk

```{r milk hk}
lf_sec_06 <- read_dta(here::here("data", "raw", "lsms","lf_sec_06.dta"))

milk <- clean_up(lf_sec_06)

lapply(lf_sec_06, attr, "label")
```

```{r milk naming}
# three households report no average; they each need to be treated differently, see below
milk <- milk %>% 
  mutate(
    # Manual fix: three households with missing average milk production (lf06_03).
    # 1001-001: only lowest value reported — use lo as average.
    # 1002-001: no average but hi and lo exist — use mean of hi and lo.
    # 2943-001: only consumption values reported — derive average from consumption sum.
    # Hardcoded by design — verified against raw data. Flag for review if raw data changes.
    lf06_03 = ifelse(y4_hhid == "1001-001", lf06_05_2, lf06_03), # only lo reported
    lf06_03 = ifelse(y4_hhid == "1002-001", (lf06_05_2+lf06_04_2)/2, lf06_03), # no av but lo and hi exist
    lf06_03 = ifelse(y4_hhid == "2943-001", lf06_07+lf06_08+lf06_10, lf06_03)) # only consumption; lists buyer despite no sales.

milk <- upData(milk,
               rename = .q(
                 lf06_01 = milked,
                 lf06_02 = length,
                 lf06_03 = av,
                 lf06_04_2 = hi,
                 lf06_05_2 = lo,
                 lf06_07 = consumed,
                 lf06_08 = sold,
                 lf06_09 = processed,
                 lf06_10 = psold,
                 lf06_11 = value,
                 lf06_12_1 = buyer1,
                 lf06_12_2 = buyer2
               ),
               
               # vars needed to determine lactation curve
               # NOTE (backlog): lactation curve estimation not implemented.
               # as.ToDate() is not a valid R function — replace with as.Date() if revived.
               # Reference: Shija et al. 2022. Low priority for MFA pipeline.
               # hi_mo = as.Date(lf06_04_1),
               # lo_mo = as.ToDate(lf06_05_1),
               
               # time between highest and lowest month -> Shija et al. 2022

               period = length * 30.437, # refers to animals as a group
               # days =  period * milked, # total days milked (all animals) - is this needed?
               milk = period*av, # milk production for the entire period animals were milked for
               milk_lo = period * lo,
               milk_hi = period*hi,
               consumed_a = consumed*period,
               sold_a = sold*period,
               processed_a = processed*period,
               psold_a = psold* period,
               output = milk/milked,
               ava = av/milked,
               value_a = value*period, # average earning for milk in produce per day from sales
               
               # improve milk quantity estimate
               # source: 
               # milk_alt = ifelse(av==hi, , milk) # estimate new average when average is the same as highest value
               # milk_alt = ifelse(av==lo, , milk) # estimate new average when average is the same as lowest value
               # milk_alt # estimate new average based on highest and lowest value
               
               labels = .q(
                 milked = "Number of animals milked in past 12 mo",
                 length = "Average number of mo animals were milked for",
                 av = "Average milk production per day per type of animal",
                 hi = "Highest milk production",
                 lo = "Lowest milk production",
                 consumed = "Quantity consumed per day",
                 sold = "Quantity sold per day",
                 processed = "Quantity of milk converted into produce per day",
                 psold = "Quantity of milk products sold per day",
                 
                 # generated vars
                 period = "Period of milking",
                 days = "Total number of days milked",
                 milk = "Total annual quantity of milk produced",
                 milk_lo = "Total estimate with lowest month",
                 milk_hi = "Total estiamte with highest month",
                 consumed_a = "Annual consumption based on milking period",
                 sold_a = "Annual sales based on milking period",
                 output = "Estimate of milk per animal",
                 ava = "Average quantity per animal"
                 
               ),
               units = .q(
                 av = l/day,
                 hi = l/day,
                 lo = l/day,
                 consumed = l/day,
                 sold = l/day,
                 processed = l/day,
                 psold = l/day,
                 milk = l,
                 consumed_a = l,
                 sold_a = l,
                 ava = l
               ))




contents(milk)

milk <- milk[lvstckcat == "large ruminants" | lvstckcat == "small ruminants"] # remove animals not milked but keep households not milking

saveRDS(milk, here::here("data", "processed", "01", "milk.RDS"), compress = T)

miss_var_summary(milk)
milk[av==hi]
milk[av==lo]
```



## Milk merged with animals owned

To check whether the information provided is plausible, milking information is cross-referenced with ownership information.

```{r milk}
animals_fin <- readRDS(here::here("data", "processed", "01", "animals_fin.RDS"))

# check number of animals owned
m <- milk[,.(y4_hhid, type = lvstckcat, milked, period, av, hi, lo, ava, milk, milk_lo, milk_hi, consumed = consumed_a, sold = sold_a, processed = processed_a, psold = psold_a)]

m[,type := as.factor(type)]

```

***The interviewer should check if the amount of milk produced (Question 3) is greater or equal to liquid milk consumed (Question7) + liquid milk sold (Question 8) + milk processed into dairy products (Question 9). Q3 has to be greater than Q7 + Q8 + Q9***



```{r}
# readjust milk estimate from average, hi, lo
# Step 1: Calculate min, max, and average from raw data
# Step 1: Adjust inconsistencies
cleaned_data <- milk %>%
  mutate(
    processed_new = ifelse(processed < psold, psold, processed), # Replace processed value
    processed_new = replace_na(processed_new, 0),
    smd1 = consumed + sold + processed_new  # Total milk used
  ) %>%
  # Calculate averages
  mutate(
    # Use smd1 as a fallback for av if av is missing
    av_with_fallback = ifelse(is.na(av), smd1, av),
    
    # Simple average of av (or smd1 as fallback), hi, and lo
    simple_av = (av_with_fallback + hi + lo) / 3,  
    
    # Weighted average using smd1 and av (fallback handled)
    weighted_av = ifelse(
      is.na(av),  # If av is missing, rely fully on smd1
      smd1,
      0.8 * av + 0.2 * smd1  # Weighted based on confidence
    ),
    
    # Geometric average including smd1 as fallback for av
    geo_av = ifelse(
      is.na(av), 
      sqrt(smd1 * smd1),  # If av is missing, use smd1
      sqrt(av * smd1)     # Otherwise, combine av and smd1
    ),
    
    # Adjust average, ensuring it falls within the bounds lo and hi
    corrected_avg = ifelse(
      is.na(av_with_fallback) | is.na(lo) | is.na(hi), 
      av_with_fallback, 
      pmin(pmax(av_with_fallback, lo), hi)
    ),
    corrected_min = ifelse(
      is.na(lo) | is.na(corrected_avg), 
      lo, 
      pmin(lo, corrected_avg)
    ),
    corrected_max = ifelse(
      is.na(hi) | is.na(corrected_avg), 
      hi, 
      pmax(hi, corrected_avg)
    ),
    
    # Step 2: Calculate uncertainty
    range = ifelse(
      is.na(corrected_max) | is.na(corrected_min), 
      NA, 
      corrected_max - corrected_min
    ),
    SD = ifelse(
      is.na(range), 
      NA, 
      range / 4  # Approximate standard deviation assuming normal distribution
    ),
    
    # Step 3: Derive a representative production value
    new_av = ifelse(
      av==smd1, av,
      # is.na(corrected_min) | is.na(corrected_avg) | is.na(corrected_max),
      # smd1,  # Use smd1 as fallback if corrected values are missing
      0.2 * corrected_min + 0.6 * corrected_avg + 0.2 * corrected_max  # Weighted adjustment
    )
  )

# cleaned_data %>% select(starts_with("corrected"), ends_with("_av"), av, hi, lo, smd1) %>% View()
  
mass_milk <- cleaned_data %>% 
  mutate_if(is.numeric, ~replace(., is.na(.), 0)) %>%
  dplyr::rename(type = lvstckcat) %>%
  mutate(
    # recalculate based on new estimates, only calculate what is needed, check for errors later
    milkwa = av * period,
    milk = new_av * period,
    SD = SD*period,
    # all annual estimates
    consumed = consumed * period,
    sold = sold * period,
    psold = psold* period,
    processed = processed * period,
    processed_new = ifelse(processed < psold, psold, processed),# replace processed (no reason to detail, quality is not getting any better)
    smd1 = consumed+sold+processed_new,
    milk = ifelse(milk==0 & smd1>0, smd1, milk),
    missing = milk-smd1# messes up SD for those households
  )
# %>% 
#   select(y4_hhid, type, milk, SD, consumed, sold, processed= processed_new, psold, missing)

saveRDS(mass_milk, here::here("data", "processed", "01", "mass_milk.RDS"), compress = T)
```


```{r}
# data checks
excl_milk <- cleaned_data %>% 
  mutate(
    new2_av = (new_av+smd1)/2,
    new3_av = (av+new_av/2)
  ) %>% 
  filter(milked>0) %>% 
  select(y4_hhid, lvstckcat, corrected_avg, corrected_min, corrected_max, ends_with("av"), av_with_fallback, SD, consumed, sold, processed, psold, period, milked) %>% 
  dplyr::rename(mean = new_av) %>% # weighted_av and geo_av give lowest exclusions
  mutate(
    processed_new = ifelse(processed < psold, psold, processed),# replace processed (no reason to detail, quality is not getting any better)
    processed_new = replace_na(processed_new, 0),
    smd1 = consumed+sold+processed_new) %>%
  mutate(
    excl = fcase(
      period>310 & milked == 1, "Implausible",
      smd1>7 & milked == 1, "Implausible", # no more than 5.5 l a day plausible with margin
      mean*1.2<consumed | mean*1.2<sold | mean*1.2<processed_new, "Data inconsistent", # exclude if production + variation are smaller than reported consumption
      # mean+(SD*2)<consumed | mean+(SD*2)<sold | mean+(SD*2)<processed_new, "Data inconsistent", # exclude if production + variation are smaller than reported consumption
      smd1 > mean * 1.5, "Excessive milk use", # as per manual with 20% margin
      smd1<= mean * 0.5, "Milk unaccounted")
  ) %>%
  mutate(
    item = paste("milk", lvstckcat, sep = " - ")
  ) %>%
  select(y4_hhid, item, excl)


# Exclusion counts by averaging method — used to select new_av (92) as best balance.
# Lower = fewer exclusions from plausible households.
# new_av chosen over weighted_av (42, too permissive) and corrected_avg (106, too conservative).
# av_with_fallback: 100
# simple_av: 104
# geo_av: 45
# weighted_av: 42, 4/6
# weighted av: 52 6/4
# weighted av: 74 8/2
# new_av 92
# corrected_av: 106
excl_milk %>% filter(!is.na(excl)) %>% nrow()
excl_milk %>% group_by(excl) %>% summarise(n = n()) %>% arrange(desc(n))

write.csv(excl_milk, here::here("data", "processed", "01", "excl_milk.csv"))
```

# Exclusions
Revise exclusions. What should the cut-off be, is it reasonable to exclude on length and quantity given we never know how many animals are milked at any point in time?


```{r milk exclusions, eval=FALSE}
# NOTE: eval=FALSE — superseded by active exclusion chunk above.
# Retained for methodological reference — used weighted average (new_av) approach.
# Delete once exclusion approach is finalised.
mcop <- m %>% 
  # duplicates code
  rowwise() %>% 
  mutate(
    processed_new = ifelse(processed < psold, psold, processed),# replace processed (no reason to detail, quality is not getting any better)
    smd1 = consumed+sold+processed_new) %>%
  mutate(
    excl = fcase(
      period>306 & milked == 1, "Implausible",
      ava>5.5 & milked == 1, "Implausible", # no more than 5.5 l a day
      av*1.5>hi, "Implausible", # average production exceeds max
      # max_owned<milked, "Implausible", # number of animals milked exceed max possibly owned possibly one...
      smd1> milk *1.5, "Excessive milk use", # as per manual with 20% margin
      smd1< milk *0.5, "Milk unaccounted",
      milk<consumed * 1.2| milk < sold * 1.2 | milk < processed_new * 1.2, "Data inconsistent"
    ),
    item = paste("milk", type, sep = " - ")
  ) 

excl_milk <- mcop %>% 
  select(y4_hhid, item, excl)

write.csv(excl_milk, here::here("data", "processed", "01", "excl_milk.csv"))
```


```{r milk exclusions old,eval=FALSE}
# NOTE: eval=FALSE — original exclusion approach using rowSums and top 2%.
# Superseded — retained for methodological reference only.
# prepare for exclusions
m$smd1 <- rowSums(m[,consumed:psold],na.rm=TRUE) # that's not very accurate, should not be the basis of decision making
m[,excl := ifelse(smd1 < milk * 0.5, "Milk unaccounted", NA)]
m[is.na(excl)]

# determine top 2% production volumes
top2 <- function(x) quantile(x, probs = .98, na.rm=TRUE) # function to calculate top 2
m[, top := top2(milk), by=type]

# top 2% of milk per animal
m[, topa := top2(ava), by=type]
# top 2 % of milking length
m[, topd := top2(period), by=type]

# not fully clear if psold should be included or is part of processed
# replace missing
# Replace NA with 0 in milk disposition columns — named to avoid positional fragility
cols_to_zero <- c("consumed", "sold", "processed")
m[, (cols_to_zero) := lapply(.SD, function(x) ifelse(is.na(x), 0, x)), .SDcols = cols_to_zero]
m[, processed_new := ifelse(processed == 0, psold, processed)] # will be dropped and recreated in MFA

# MILK ASSUMPTIONS
m$smd1 <- rowSums(m[,consumed:processed],na.rm=TRUE) # as per interviewer manual
m$smd2 <- rowSums(m[,consumed:psold],na.rm=TRUE) # processed is not accurate, therefore extend to psold
# replace processed
# given processed and psold is inaccurate a decision to exclude on this basis may not be justifiable
m[, smd3 := consumed+sold+processed_new]

# spread of inaccuracy
m[, diff :=  round(milk - smd3,2)]
m[, diff := ifelse(diff<0, diff*-1, diff)]
m[, perc := diff*100/milk]

# exclude implausible number of animals milked
ac <- animals_fin[,.(y4_hhid, type, sex, current, max_owned)]

# collapse to household level
ac <- ac[type=="small ruminants" | (type == "large ruminants" & sex =="female")]
ac <- ac[, lapply(.SD, sm), .SDcols=is.numeric, by=.(y4_hhid, type)]

m <- ac[m, on=c("y4_hhid", "type")] 

m[, milked := as.numeric(milked)]

# exclusions
m <- m[, excl := fcase(
    # av == hi,   "Av==hi", 
    # av == lo, "Av==lo", 
    # smd2 > milk * 1.2, "Milk insufficient", # household used more than they produced
    # smd2 < milk * 0.8, "Milk unaccounted", # household produced more than they used
    # top <= milk, "Top 2%", # not sure this is valid as we do not know the number of animals used
    # topd <= period, "Top 2% length", # length is rarely for an individual animal which makes in theory all length possible 
    period>306 & milked == 1, "Implausible",
    max_owned<milked, "Implausible animal number", # number of animals milked exceed max possibly owned
    perc > 100, "Quantities mismatch", # gross mismatch of quantities produced and used
    smd1> milk *1.5, "Quantities mismatch" # as per manual with 20% margin
    # topa <= ava, "Top 2% per animal", # impossible to determine
    # period > 305, "Implausible milking length" # length exceeds plausible check refs! might to apply to the entire group
    )]

# work around, should be integrated in fcase code below
ex <- m %>% 
  filter(is.na(excl)) %>% 
  filter_at(vars(consumed:psold), any_vars(. > milk*1.2))  %>% 
  mutate(excl = "Data inconsistent")

# remove ex from c and add back in
m <- m[!ex, on = .(y4_hhid, type)]
excl_milk <- bind_rows(m, ex)

excl_milk <- excl_milk %>% 
  mutate(item = paste("milk", type, sep = "-")) %>% 
  select(y4_hhid, item, excl)
# m[,excl:= as.factor(excl)]

write.csv(excl_milk, here::here("data", "processed", "01", "excl_milk.csv"))
```

## Milk mass

```{r}
milk2 <- m %>% 
  mutate_if(is.numeric, ~replace(., is.na(.), 0)) %>% # should be full list of livestock owning households
  rowwise(y4_hhid) %>% 
  mutate(
    milk_mean = (milk+milk_lo+milk_hi)/3,# generate mean production and uncertainty (what does this mean for feed)
    # generate uncertain and include in mean estimate for production
    processed_new = ifelse(processed < psold, psold, processed),# replace processed (no reason to detail, quality is not getting any better)
    smd1 = consumed+sold+processed_new
  ) %>% 
  select(y4_hhid, type, milk_mean, milk, milk_hi, milk_lo, consumed, sold, processed = processed_new, psold, smd1)


milk3 <- upData(milk2,
                   rename = .q(
                     smd1 = sm_disp
                   ),
                   labels = .q(
                     processed_new = "Stand-in: product sold equals input bc missing",
                     sm_disp = "Sum of all disposition",
                     processed = "Annual quantity used for processing",
                     sold = "Annual quantity sold",
                     psold = "Quantity of product sold",
                     mean_milk = "Average across lowest, highest and average milk estimate"
                   ), 
                   drop = .q(need))

setDT(milk3)
```


Aim:

-   Estimate feed consumed for milk produced

```{r milk feed}
# feed information from animals file
f <- readRDS(here::here("data", "processed", "01", "feed_short.RDS"))

# @Opio.2013 p.119 "Small ruminant milk feed ration"
smrum <- data.table(
  feed1 = c("only feeding (no grazing/scavenging)",
            "mainly grazing/scavenging w/ some feeding",
            "only grazing/scavenging",
            "mainly feeding w/ some grazing/scavenging",
            "tethering"),
  feed = c(1, 0.35, 0, 0.65, 0.5),
  grazed = c(0, 0.65, 1, 0.35, 0.5),
  type = "small ruminants")

# @Opio.2013 p.117 dairy cattle
lgrum <-   data.table(
  feed1 = c("only feeding (no grazing/scavenging)",
            "mainly grazing/scavenging w/ some feeding",
            "only grazing/scavenging",
            "mainly feeding w/ some grazing/scavenging",
            "tethering"),
  feed = c(1,0.25,0,0.75, 0.5),
  grazed = c(0,0.75,1,0.25, 0.5), 
  # hay + fresh grass p.117
  type = "large ruminants")

# bind rows
milkfeed <- bind_rows(smrum, lgrum)

# select only vars needed
# mf <- m[,.(y4_hhid, type, milk, consumed, sold, processed, psold, smd1)]

# merge with feed
mf <- f[milk3, on=c("y4_hhid", "type")] 

# merge with feed conversion
mf <- milkfeed[mf, on=c("type", "feed1")]

# calculate feed required
mf <- upData(mf,
             need = milk * 0.7, # currently no difference between small & large ruminants, based on @Alexander.2016
             feed = need*feed,
             grazed = need*grazed,
             
             labels = .q(
               feed = "Quantity of feed consumed in DM",
               grazed = "Quantity grazed/scavenged in DM",
               need = "Quantity DM needed to produce"
             ),
             
             units = .q(
               feed = "kg DM",
               grazed = "kg DM",
               need = "kg DM",
               milk = kg,
               consumed = kg, 
               sold = kg, 
               processed = kg, 
               psold = kg
             ),
             drop = .q(feed1)
             )

# final mass milk
mass_milk_old <- mf %>% 
   mutate(across(milk_mean:psold, ~.x * 1.08)) # convert all into kg

saveRDS(mass_milk_old, here::here("data", "processed", "01", "mass_milk_old.RDS"), compress = T)
# NOTE (backlog): naming inconsistency — mass_milk_old is the feed-augmented final output,
# mass_milk is the intermediate. Rename to mass_milk_interim / mass_milk_final
# once downstream dependencies in assembly script are confirmed.
```



