---
title: "04_Destinations"
author: "Heike Rolker"
date: "2024-01-10"
output: html_document
---

```{r setup, include=FALSE}
library(here)
source(here::here("scripts", "packages.R"))
source(here::here("scripts", "functions.R"))
```

Overview:

This file merges production information and the destination of the production. It estimates the uncertainty of these quantities.

Final files:

- **tree_disp/crop_disp** mostly raw, cleaned files



- **chh/thh** (not saved) crop production at household level (ignoring plot info)

# Crop disposition

Final:

- production, dispositions & uncertainty

- quantities sold & buyers

- exclusions: y4_hhid, item, excl

```{r disp crops hk}
# load data
ag_sec_5a <- read_dta(here::here("data", "raw", "ag_sec_5a.dta")) # --> long
ag_sec_5b <- read_dta(here::here("data", "raw", "ag_sec_5b.dta")) # --> short

# prep for join
long <- prep(ag_sec_5a, season = "long")
short <- ag_sec_5b %>% 
  prep("short") %>% 
  strip_colnames("5b", "5a")

# extract labels (short contains one additional variable)
labs <- prep_labs(short)

# join & clean

disp <- long %>% 
  bind_dt(short)
setnames(disp,"zaocode","cropid") 
disp <- disp %>% clean_names(list = crops_list, "cropid")

label(disp) <- as.list(labs[match(names(disp), names(labs))])

contents(disp)
```

```{r crop disp naming}
head(disp)
crop_disp <- upData(disp,
               rename = .q(
                 ag5a_01 = sale,
                 ag5a_02 = sold,
                 ag5a_03 = value_sale,
                 ag5a_04 = b1,
                 ag5a_05 = soldb1,
                 ag5a_06 = value_b1,
                 ag5a_08 = n_int1,
                 ag5a_11 = b2,
                 ag5a_12 = soldb2,
                 ag5a_13 = value_b2,
                 ag5a_15 = n_int2,
                 ag5a_23 = storage,
                 ag5a_24 = stored,
                 ag5a_28 = storage_purpose,
                 ag5a_29 = lost,
                 ag5a_30 = loss_cause,
                 ag5a_32 = value_losses,
                 ag5a_32a = consumed,
                 ag5a_32b = seed,
                 ag5a_32c = payment,
                 ag5a_32d = gifts,
                 ag5a_32e = feed,
                 ag5a_33 = residue_use,
                 ag5a_34 = residue,
                 ag5a_35 = value_residue
               ),
               
               losses = ag5a_31/10,
               
               # replace NA with 0 where true 
               sold = ifelse(sale == "no", 0, sold),
               soldb1 = ifelse(sale == "no", 0, soldb1),
               soldb2 = ifelse(sale == "no", 0, soldb2),
               stored = ifelse(storage == "no", 0, stored),
               losses = ifelse(lost == "no", 0, losses),
               
               labels = .q(
                      sale = "Item was sold",
                      sold = "Quantity sold",
                      soldb1 = "Quantity sold to main buyer",
                      soldb2 = "Quantity sold to second largest buyer",
                      storage = "Item is being stored",
                      stored = "Quantity of item in storage",
                      losses = "Proporition of harvest lost post-harvest",
                      loss_cause = "Cause of ph losses",
                      consumed = "Quantity of item consumed",
                      seed = "Quantity stored for seed",
                      payment = "Quantity of item given as payment",
                      gifts = "Quantity of item given as gifts",
                      feed = "Quantity of item used as animal feed",
                      residue = "Quantity of residue sold"
                    ),
               units = .q(
                 sold = kg,
                 value_sales = 'T shilling',
                 soldb1 = kg,
                 value_b1 = 'T shilling',
                 soldb2 = kg,
                 value_b2 = 'T shilling',
                 storage = kg,
                 losses = percentage,
                 consumed = kg,
                 seed = kg,
                 payment = kg,
                 gifts = kg,
                 feed = kg,
                 residue = kg,
                 value_residue = 'T shilling'
               ))

# Manual fix: household 8659-001 has a clear data entry error for maize consumption.
# Hardcoded by design — verified against raw data. Flag for review if raw data changes.
crop_disp[y4_hhid == "8659-001" & cropid == "maize", consumed := 480] # clear typo
saveRDS(crop_disp, file = here::here("data", "processed", "01", "crop_disp.RDS"), compress = TRUE)
```

Transport information could allow transport

```{r disp stats}
miss_var_summary(crop_disp)

crop_disp[is.na(losses)]
```


# Tree disposition

Final:

- production, dispositions & uncertainty

- quantities sold & buyers

- exclusions: y4_hhid, item, excl

```{r tree disp hk}
# same as for crops disp
ag_sec_7a <- read_dta(here::here("data", "raw", "ag_sec_7a.dta")) # --> fruit
ag_sec_7b <- read_dta(here::here("data", "raw", "ag_sec_7b.dta")) # --> permanent

# prep for join
fruit <- prep(ag_sec_7a, season = "fruit")
perm <- ag_sec_7b %>% 
  prep("permanent") %>% 
  strip_colnames("7b", "7a")

# extract labels (short contains one additional variable)
labs <- prep_labs(perm)

# join & clean
disp <- fruit %>% 
  bind_dt(perm)

setnames(disp,"zaocode","cropid") 
disp <- disp %>% clean_names(list = crops_list, "cropid")


label(disp) <- as.list(labs[match(names(disp), names(labs))])

contents(disp)
```

```{r tree disp naming}
tree_disp <- upData(disp,
                    rename = .q(
                      ag7a_02 = sale,
                      ag7a_03 = sold,
                      ag7a_04 = value_sale,
                      ag7a_07_1 = b1,
                      ag7a_07_2 = b2,
                      ag7a_08 = storage,
                      ag7a_09 = stored,
                      ag7a_13 = lost,
                      ag7a_14 = loss_cause,
                      ag7a_16 = value_losses,
                      ag7a_17 = consumed,
                      ag7a_18 = seed,
                      ag7a_19 = payment,
                      ag7a_20 = gifts,
                      ag7a_21 = feed
                    ),
                    
                    losses = ag7a_15/10,
                    
                    # replace NA with 0 where true 
                    sold = ifelse(sale == "no", 0, sold),
                    stored = ifelse(storage == "no", 0, stored),
                    losses = ifelse(lost == "no", 0, losses),
                    
                    labels = .q(
                      sale = "Item was sold",
                      sold = "Quantity sold",
                      b1 = "Buyer 1",
                      b2 = "Buyer 2",
                      storage = "Item is being stored",
                      stored = "Quantity of item in storage",
                      losses = "Proporition of harvest lost post-harvest",
                      loss_cause = "Cause of ph losses",
                      consumed = "Quantity of item consumed",
                      losses = "Fraction of harvest lost",
                      seed = "Quantity stored for seed",
                      payment = "Quantity of item given as payment",
                      gifts = "Quantity of item given as gifts",
                      feed = "Quantity of item used as animal feed",
                      residue = "Quantity residue sold"
                    ),
                    units = .q(
                      sold = kg,
                      value_sale = 'T Shilling',
                      stored = kg,
                      consumed = kg,
                      losses = percentage,
                      seed = kg,
                      payment = kg,
                      gifts = kg,
                      feed = kg
                    ))

saveRDS(tree_disp, file = here::here("data", "processed", "01", "tree_disp.RDS"), compress = TRUE)
```


## Merge production with disposition

Aim:

-   determine annual production for each household 

```{r crops hh}
# crop/plot info
pc <- readRDS(here::here("data", "processed", "01", "pc.RDS")) # to keep code intact

# number of plots per crop and household
pn <- pc %>% 
  select(y4_hhid, plotnum, cropid) %>% 
  count(y4_hhid, cropid) %>% 
  dplyr::rename(nplots = n) 

pm <- pc %>% 
  group_by(y4_hhid, cropid) %>% 
  summarise(
    mismatch = sum(mismatch)
  ) %>% 
  ungroup() 

pn <- merge(pn, pm, by = c("y4_hhid", "cropid"), all=TRUE)

# necessary?
chh <- copy(pc)

# drop vars not needed
chh[, .q(
    area_planted, # converted to absolute value
    gps_area_new, # replaced by plotsize
    area_new, # replaced by plotsize
    harvested, # used to determine remaining harvest
    lessharvest, # used to determine area harvested
    harvest_remain, # converted to absolute value
    area_harvested_new # too. empty to be useful
  ) := NULL]

# bring production to household level
# check if this code is correct and doesnt keep duplicates --> looks like it's all fine
chh <- pc[,.(harvest = sum(quant_harvest),
          quant_unharvested = sum(quant_unharvested),    
          total_harvest = sum(total_harvest),
          area_planted = sum(area_planted_new)), # no need to keep the "new"
          by=.(y4_hhid, type, cropid)] # group by hh, type and cropid

# estimate kg/ha 
chh[, yield := total_harvest/area_planted]

# add plotnumbers back in --> not needed, keep for code
chh <- merge(chh, pn, by=c("y4_hhid", "cropid"), all=TRUE)

pn[mismatch!=0]
unique(chh)
```

-   determine destination of all production

```{r crops destination}
cd <- copy(crop_disp)

# select disposition vars
cd <- cd[, .(y4_hhid, type, cropid,
                    sold, stored, losses, consumed, seed, payment, gifts, feed, residue)]

# collapse disposition to household level (ignores the seasons)
cd <- cd[, lapply(.SD, sm), .SDcols=is.numeric, by=.(y4_hhid, type, cropid)]

crops <- merge(chh, cd, by=c("y4_hhid", "type", "cropid"), all=TRUE)

# quantify losses
crops <- upData(crops,
                losses = losses * harvest,
                labels = .q(
                      sold = "Quantity sold",
                      stored = "Quantity of item in storage",
                      consumed = "Quantity of item consumed",
                      seed = "Quantity stored for seed",
                      payment = "Quantity of item given as payment",
                      gifts = "Quantity of item given as gifts",
                      feed = "Quantity of item used as animal feed",
                      losses = "Quantity lost ph",
                      residue = "Quantity residue sold",
                      harvest = "Annual quantity harvested",
                      nplots = "Number of plots crop planted on",
                      mismatch = "Number of plots where plotsize smaller than area harvested"
                      ),
                units = .q(
                      sold = kg,
                      stored = kg,
                      consumed = kg,
                      losses = kg,
                      seed = kg,
                      payment = kg,
                      gifts = kg,
                      feed = kg,
                      residue = kg
                    ),
                drop = .q(
                  area_planted, 
                  total_harvest,
                  nplots,
                  mismatch # no need
                ))

# add sum for all disposition
#crops$smd <- rowSums(crops[,sold:residue],na.rm=TRUE) # residue not part of production volume, seed should also not be included
crops[, smd:= sold+stored+losses+consumed+payment+gifts+feed]
crops[, smd := adlab(smd, 'Sum of all disposition')]

# overview
miss_var_summary(crops)
contents(crops)

# save
saveRDS(crops, here::here("data", "processed", "01", "mass_crops.RDS"), compress = TRUE)
```

-   Full-merge & exclude:

    -   top 2% of household production for each crop (crop failure and bottom 2% is more plausible)

```{r exclusions crops}
crops_excl <- copy(crops)
# NOTE (backlog): top2() already defined in functions.R — local definition redundant.
# Remove once confirmed consistent with functions.R definition across all scripts.
top2 <- function(x) quantile(x, probs = .99, na.rm=TRUE) # function to calculate top 2
crops_excl[, top := top2(yield), by=cropid]

# add number of occurances -> for what?
cropnums <- crops_excl %>% group_by(cropid) %>% summarise(nums = n()) %>% setDT()
# merge back in -> find more elegant solution
crops_excl <- cropnums[crops_excl, on = .(cropid)]

crops_excl[, yield := as.numeric(yield)] # to avoid class error below
# too big to process...
# c %>%
#   filter(nums >10) %>%
#   ggplot( aes(x=yield, color=cropid, fill=cropid)) +
#     geom_histogram(alpha=0.6, binwidth = 5) +
#     # scale_fill_viridis(discrete=TRUE) +
#     # scale_color_viridis(discrete=TRUE) +
#     # theme_ipsum() +
#     # theme(
#     #   legend.position="none",
#     #   panel.spacing = unit(0.1, "lines"),
#     #   strip.text.x = element_text(size = 8)
#     # ) +
#     facet_wrap(~cropid)

# c %>% filter(cropid == "cucumber") %>% 
#   ggplot(aes(yield)) + geom_histogram()

crops_excl[, diff := harvest - smd]
crops_excl[, diffp := diff*100/harvest]

# more elegant solution
# below could have a more explicit hierarcy by adding & is.na(excl)
crops_excl <- crops_excl[, excl := fcase(
    is.na(sold),   "Data missing",
    is.na(yield), "Data missing",
    # mismatch != 0, "Harvested area exceeds plotsize", # overly strict and not too relevant for my purposes
    # top <= yield & nums >10, "Top 2%" # this can only be done for these crops here not for the permanent/tree crops, only keeping exclusion criteria that matter to internal validity and consistency
    smd > harvest * 1.3, "Harvest insufficient",
    smd < harvest * 0.7, "Harvest unaccounted"
                  )]

# work around, should be integrated in fcase code below
ex <- crops_excl %>% 
  clear.labels() %>% 
  filter(is.na(excl)) %>% 
  filter_at(vars(sold:residue), any_vars(. > harvest*1.2)) %>% 
  mutate(excl = "Data inconsistent")

# remove ex from c and add back in
crops_excl <- crops_excl[!ex, on = .(y4_hhid, cropid)]
crops_excl <- clear.labels(crops_excl)
excl_crops <- bind_rows(crops_excl, ex)

miss_var_summary(crops_excl)

crops_excl[!is.na(excl)]
# View(c[top <= yield])

excl_crops <- excl_crops %>% 
  select(y4_hhid, item = cropid, excl)

# save list of exclusions
write.csv(excl_crops, here::here("data", "processed", "01", "excl_crops.csv"))
```

#### Trees

Aim:

-   determine annual production volumes

-   calculate yield (i.e., harvest/tree) --\> not a useful metric

```{r trees hh}
pt <- readRDS(here::here("data", "processed", "01", "pt.RDS"))

# number of plots per crop and household
pn <- pt %>% 
  select(y4_hhid, plotnum, cropid) %>% 
  count(y4_hhid, cropid) %>% 
  dplyr::rename(nplots = n)

# necessary
thh <- copy(pt)

# bring production to household level
thh <- thh[,.(harvest = sum(harvest), ntrees = sum(ntrees)),
          by=.(y4_hhid, type, cropid)]

# calculate yield
thh[, yield := harvest/ntrees]

# add plotnumbers back in
thh <- merge(thh, pn, by=c("y4_hhid", "cropid"), all=TRUE)

miss_var_summary(thh)
# thh[is.na(yield)] # remove all
# thh[is.na(harvest)] # goes with yield
# thh[is.na(ntrees)] # goes with yield
```

```{r trees destinations}
# read in data
tree_disp <- readRDS(here::here("data", "processed", "01", "tree_disp.RDS"))
td <- copy(tree_disp)

# select disposition vars
td <- td[, .(y4_hhid, type, cropid,
                    sold, stored, losses, consumed, seed, payment, gifts, feed)]

# collapse disposition to household level (ignores the seasons)
td <- td[, lapply(.SD, sm), .SDcols=is.numeric, by=.(y4_hhid, type, cropid)]

# merge trees and disposition
trees <- merge(thh, td, by=c("y4_hhid", "type", "cropid"), all=TRUE)

# quantify losses
trees <- upData(trees,
                losses = losses * harvest,
                labels = .q(
                      sold = "Quantity sold",
                      stored = "Quantity of item in storage",
                      consumed = "Quantity of item consumed",
                      seed = "Quantity stored for seed",
                      payment = "Quantity of item given as payment",
                      gifts = "Quantity of item given as gifts",
                      feed = "Quantity of item used as animal feed",
                      losses = "Quantity lost ph",
                      # residue = "Quantity residue sold",
                      harvest = "Annual quantity harvested",
                      # nplots = "Number of plots crop planted on"
                      yield = "Yield (kg) per tree"
                      ),
                units = .q(
                      sold = kg,
                      stored = kg,
                      consumed = kg,
                      losses = kg,
                      seed = kg,
                      payment = kg,
                      gifts = kg,
                      feed = kg,
                    ))

# sum of disposition
# trees$smd <- rowSums(trees[,sold:feed], na.rm=TRUE) 
trees[, smd := sold+stored+losses+consumed+payment+gifts+feed]
trees[, smd := adlab(smd, 'Sum of all disposition')]

miss_var_summary(trees)
# trees[is.na(yield)] # scattered info, one entry might be worth saving
# trees[is.na(sold)] # no disposition info

saveRDS(trees, here::here("data", "processed", "01", "mass_trees.RDS"), compress = T)
```

```{r exclusions trees}
trees_excl <- copy(trees)

trees_excl[, top := top2(yield), by=cropid]

trees_excl[, `:=` (harvest = as.numeric(harvest),
          ntrees = as.numeric(ntrees))] # same class problem as above

# more elegant solution
trees_excl <- trees_excl[, excl := fcase(
    is.na(sold),   "Incomplete", # no disposition info
    is.na(yield), "Incomplete", 
    ntrees = 0 & harvest > 0, "Non-match", # numbers do not match up
    # top <= yield, "Top 2%",
    smd > harvest * 1.3, "Harvest insufficient",
    smd < harvest * 0.7, "Harvest unaccounted"
                  )]

# work around, should be integrated in fcase code below
ex <- trees_excl %>% 
  clear.labels() %>% 
  filter(is.na(excl)) %>% 
  filter_at(vars(sold:feed), any_vars(. > harvest*1.2)) %>% 
  mutate(excl = "Data inconsistent")

# remove ex from c and add back in
trees_excl <- trees_excl[!ex, on = .(y4_hhid, cropid)]

check <- trees_excl[is.na(excl)]
miss_var_summary(check)

check %>% 
  rowwise() %>%
  filter_at(vars(sold:feed), any_vars(. > harvest*1.1)) # affects multiple pre-exclusion and 1 obs post -> ignore

excl_trees <- trees_excl %>% 
  clear.labels() %>% 
  bind_rows(ex) %>% 
  select(y4_hhid, item = cropid, excl)

write.csv(excl_trees, here::here("data", "processed", "01", "excl_trees.csv"))
```


## Crops & crops


Aim:

-   align information for crops and permanent crops to only reflect the overlap

| Crop | Quantity | Processed   | *Product* | Consumed | Sold | *Destination* | Unaccounted, lost or wasted |
|---------|---------|---------|---------|---------|---------|---------|---------|
|      |          | sum of all? |           |          |      |               |                             |
|      |          |             |           |          |      |               |                             |
|      |          |             |           |          |      |               |                             |

```{r crops trees}
# select vars
ct <- crops[, .(y4_hhid, type, cropid, harvest, sold, stored, seed, losses, consumed, payment, gifts, feed, residue, smd)]
# ct[, source:= "Crops"] # possibly for the purpose of consort

tc <- trees[, .(y4_hhid, type, cropid, harvest, sold, stored, seed, losses, consumed, payment, gifts, feed, smd)] 

tc[, residue := NA] # no residue in trees data
# tc[, source := "Trees"] # possibly for the purpose of consort

# match vars --> unelegant work-around...
tc <- upData(tc,
             harvest = as.integer(harvest),
             residue = as.integer(residue),
             labels = .q(
               residue = "Quantity residue sold",
               harvest = "Annual quantity harvested"
             ),
             units = .q(
               residue = kg,
               harvest = kg))

# bind rows
allcrops <- rbindlist(list(ct, tc), fill = TRUE)

# split sm --> move to balancing in MFA
# allcrops <- upData(allcrops,
#                    stock = ifelse(harvest-smd<0, (harvest-smd)*(-1), NA), # assumes excess disposition are stocks from previous year
#                    missing = ifelse(harvest-smd>0, harvest-smd, NA), # assumes missing,
#                    labels = .q(
#                      stock = "Neg difference between harvest and disposition",
#                      missing = "Positive difference between harvest and disposition"
#                    ))

# allcrops[, lapply(.SD, sm), .SDcols=is.numeric, by=.(y4_hhid, type, cropid)] # no duplication in crops & trees 

# include quantities processed (excluded from sum variable based on assumptions)

saveRDS(allcrops, file = here::here("data", "processed", "01", "mass_allcrops.RDS"), compress = TRUE)
```

Bonus:

-   include volumes sold across destinations

    *Destination will only ever cover recipients/buyers partially as only two main buyers are listed. It could be of interest to see whether they cover all/most of the production, i.e., to understand how divers the sale channels are (CHAPTER 4?) Tree crops do not quantify how much goes to certain buyers.*

### Crops & products

Aim:

-   collapse all processed to one quantity and output for merge with production info -> already done in ag_produce file
-   disentangle inputs (joint and separate accounts to avoid double counting) -> not necessary anymore

```{r, eval=FALSE}
# NOTE: eval=FALSE — agprod_mass_fin not yet finalised upstream.
# Re-enable once ag produce assembly is complete and agprod_mass_fin.RDS is stable.
agprod_mass_fin <- readRDS(here::here("data", "processed", "01", "agprod_mass_fin.RDS"))
agprod_mass_fin[, cropid := adlab(cropid, '')]

allcrops_mass <- merge(allcrops, agprod_mass_fin, by=c("y4_hhid", "cropid"), all=TRUE)
```

Exclusions

```{r complete crops, eval=FALSE}
# NOTE: eval=FALSE — exclusion logic for combined crops+products not finalised.
# Positional column indexing present (fragile) — replace with named columns before enabling.
cropscompl <- copy(allcrops_mass)

# exclude produce
# cropscompl[, excl := ifelse(input<produced, "Implausible", NA)] # step already taken in excl_agprod
cropscompl[, excl := ifelse(harvest*1.5<produced, "Implausible processing to harvest", NA)] # most reliable exclusion I can make (should be based on input but input is estimated through disentanglement), difference between input and product exclusion is 4ish

# replace NAs
# cropscompl[, 4:14][is.na(cropscompl[, 4:14])] <- 0 
# cropscompl[, 16:23][is.na(cropscompl[, 16:23])] <- 0 

excl_allcrops <- cropscompl %>% 
  select(y4_hhid, item = cropid, excl)

# miss_var_summary(cropscompl)
write.csv(excl_allcrops, here::here("data", "processed", "01", "excl_allcrops.csv"))
```

# Agricultural produce exclusions


Join with production information to determine exclusions.


```{r exclusions}
inputs_final <- readRDS(here::here("data", "processed", "01", "mass_agprod_new.RDS"))
# join with production information
mass_allcrops <- readRDS(here::here("data", "processed", "01", "mass_allcrops.RDS")) %>% 
  clear.labels()

inputs_updated <- inputs_final %>%   # estimate destination of produce
  clear.labels() %>% 
  left_join(select(mass_allcrops,y4_hhid, cropid, harvest, sold_raw = sold, consumed_raw = consumed, stored, smd), by =c("y4_hhid", "cropid")) %>%
  rowwise() %>% 
  mutate(diff = harvest-smd, # unallocated production
         vali = sold_raw+consumed_raw>=input, # true if processing fits into consumption and sales
         needed = ifelse(vali == FALSE, input - sold_raw - consumed_raw, 0),
         newprocessing = ifelse(diff>=needed, needed, diff), # processing that can be allocated
         balance = harvest-smd-newprocessing, # difference between harvest and sum of all disposition
         smd2 = smd+newprocessing, # new sum of all disposition
         overrep = harvest-smd-needed) 
# , # processing that needs to be allocated
#          diffsold = sold_raw-input_sales, # raw minus processing, negative when processing is more than raw
#          diffcons = consumed_raw-input_cons, 
#          availableoverall = ifelse(vali == FALSE, diff-input, 0), # available for processing beyond sold and consumed, negative are overreported
#          available = fcase( # available within sold and consumed
#            diffcons>0 & diffsold <= 0, diffcons,
#            diffsold>0 & diffcons <= 0, diffsold,
#            diffsold>0 & diffcons>0, diffsold+diffcons,
#            TRUE, 0
#          )),
#          voltofind = ifelse(required<0, required - available, 0), # processing that still needs to be allocated
#          newprocessing = ifelse(diff>0, ) %>% # difference between harvest and sum of all disposition, still has vol available
  
saveRDS(inputs_updated, here::here("data", "processed", "01", "mass_agprod_newnew.RDS"), compress = T)
# NOTE (backlog): filename mass_agprod_newnew is a versioning artefact.
# Rename to mass_agprod_final once downstream dependencies are confirmed.
  
excl_agprod <- inputs_updated %>% 
  mutate(excl = ifelse(produced>input, "Processing data inconsistent", NA),
         excl = ifelse(smd2>harvest*1.3, "Data inconsistent", excl),
         excl = ifelse(smd < harvest*0.7, "Data inconsistent", excl), # harvest is too small to account for all disposition
         # compare products to unallocated
         # compare products to consumption/sales quantities
         # excl = ifelse(input_cons > consumed_raw * 3 & input_cons > 0, "Consumption mismatch", excl), # we do not know whether this is/should be included in the consumption count, should it be among the "missing" values?
         # excl = ifelse(input_sales > sold_raw * 3 & input_sales > 0, "Sales mismatch", excl),
         # excl = ifelse(input>harvest * 1.5, "More processed than harvested", excl)
    # excl = ifelse(input_cons > consumed_raw + stored & input_cons > 0, "Consumption mismatch", NA), # possibly exclude...
    # excl = ifelse(input_sales > sold_raw + stored & input_sales >0, "Sales mismatch", excl)
  ) 

# Diagnostic: checking whether processing inputs are accounted for within raw
# consumption/sales. High counts suggest inputs generally sit inside consumption —
# supports current exclusion logic. Retained for methodological reference.
excl_agprod %>% filter(consumed > 0 & consumed_raw<input_cons) # 153
excl_agprod %>% filter(consumed > 0 & consumed_raw<input_cons & input_cons <= diff)
excl_agprod %>% filter(consumed > 0 & consumed_raw>input_cons) # 1008
# -> suggests input is generally accounted for inside consumption accounting
excl_agprod %>% filter(sold > 0 & sold_raw<input_sales) # 36
excl_agprod %>% filter(sold > 0 & sold_raw>input_sales) # 9
# -> suggests input is generally accounted for outside sales accounting

table(excl_agprod$excl)

excl_agprod <- excl_agprod %>%
   select(y4_hhid, item = cropid, excl)

write.csv(excl_agprod, here::here("data", "processed", "01", "excl_agprod.csv"))
```


# Livestock network

- number and weight slaughtered and destinations

```{r network hk}
# load data
lf_network <- read_dta(here::here("data", "raw", "lf_network.dta"))

# NOTE: clean_up() not used here — lfnr_02 requires manual integer decode before
# factor conversion. Manual tolower() applied to lfnr_03 to match clean_up() behaviour.
lf_network <- lf_network %>% 
  # part of clean up
  setDT() %>% 
  # mutate_all(na_if, "") %>% 
  mutate(
    lfnr_03 = tolower(as_factor(lfnr_03)),
    lfnr_02 = tolower(case_when(
      lfnr_02 == 1 ~ "RELATIVE",
      lfnr_02 == 2 ~ "NEIGHBOR",
      lfnr_02 == 3 ~ "FRIEND",
      lfnr_02 == 4 ~ "MARKET",
      lfnr_02 == 5 ~ "OPEN MARKET",
      lfnr_02 == 6 ~ "COOPERATIVE UNION",
      lfnr_02 == 7 ~ "FARMERS PARTY",
      lfnr_02 == 8 ~ "MAIN PLOT/FARM",
      lfnr_02 == 9 ~ "PRIVATE BUSINESS PERSON",
      lfnr_02 == 10 ~ "MAIN MARKET",
      lfnr_02 == 11 ~ "BUSINESS CONTACT",
      lfnr_02 == 12 ~ "ABATTORI/FACTORY",
      lfnr_02 == 13 ~ "EMPLOYER",
      lfnr_02 == 14 ~ "RELIGIOUS INSTITUTION",
      lfnr_02 == 15 ~ "BANK (COMMERCIAL)",
      lfnr_02 == 16 ~ "GROCERY/LOCAL MERCHANT",
      lfnr_02 == 17 ~ "MONEY LENDER",
      lfnr_02 == 18 ~ "NGO",
      lfnr_02 == 19 ~ "DISTRIBUTION OFFICER")),
    lfnr_id = na_if(lfnr_id, "")) %>% 
  
  # remove confidential & duplicated vars
  select(!c(lfnr_01, occ))

saveRDS(lf_network, here::here("data", "processed", "01", "lf_network.RDS"), compress = TRUE)
```

## Animal buyers

```{r animal buyers}
# slaugter information
mass_animals <- readRDS(here::here("data", "processed", "01", "mass_animals.RDS"))
# ownership and buyer information
animals <- readRDS(here::here("data", "processed", "01", "animals.RDS"))

animals <- animals[,.(y4_hhid, type, lvstckid, gifted, recp, sold, dest1, dest2, slaughter, ssold, b1 = dest1s, b2 = dest2s)]

# merge first buyer of animals slaughtered
lf_net <- lf_network[,.(y4_hhid, b1 = lfnr_id, b1_rel = lfnr_02, b1_loc = lfnr_03)]
animal_sales <- lf_net[animals, on = .(y4_hhid, b1)]

# merge second buyer of animals slaughtered
lf_net <- lf_network[,.(y4_hhid, b2 = lfnr_id, b2_rel = lfnr_02, b2_loc = lfnr_03)]
animal_sales <- lf_net[animal_sales, on = .(y4_hhid, b2)]

# merge frist buyers of animals sold alive
lf_net <- lf_network[,.(y4_hhid, dest1 = lfnr_id, dest1_rel = lfnr_02, dest1_loc = lfnr_03)]
animal_sales <- lf_net[animal_sales, on = .(y4_hhid, dest1)]

# merge second buyer of animals sold alive
lf_net <- lf_network[,.(y4_hhid, dest2 = lfnr_id, dest2_rel = lfnr_02, dest2_loc = lfnr_03)]
animal_sales <- lf_net[animal_sales, on = .(y4_hhid, dest2)]

# final quantities & weight of animals passed on
asold <- mass_animals[,.(y4_hhid, lvstckid, sold_weight)]
buyer_animals <- asold[animal_sales, on = .(y4_hhid, lvstckid)]

buyers_animals <- buyer_animals %>% 
  select(y4_hhid, type, lvstckid, sold_alive = sold, dest1_rel, dest1_loc, dest2_rel, dest2_loc, gifted, recp, slaughter, ssold, sold_weight, b1, b1_rel, b1_loc) # b2 is empty

saveRDS(buyers_animals, here::here("data", "processed", "01", "buyers_animals.RDS"), compress = T)
```

Aim: → identify flows of animals given and received

-   link livestock info to buyers and locations

-   buyers of living and dead animals, recipients of gifts and payments

-   donors of living animals



## Milk buyers

Aim:

-   show major buyers of milk produce

```{r milk buyers}
# buyer information
milk <- readRDS(here::here("data", "processed", "01", "milk.RDS"))
# milk quantities
milkfin <- readRDS(here::here("data", "processed", "01", "mass_milk.RDS"))

milk_sales <- milk[,.(y4_hhid, lvstckcat, sold, b1 = buyer1, b2 = buyer2)]

# merge first buyer
lf_net <- lf_network[,.(y4_hhid, b1 = lfnr_id, b1_rel = lfnr_02, b1_loc = lfnr_03)]
milk_sales <- lf_net[milk_sales, on = .(y4_hhid, b1)]

# merge second buyer
lf_net <- lf_network[,.(y4_hhid, b2 = lfnr_id, b2_rel = lfnr_02, b2_loc = lfnr_03)]
milk_sales <- lf_net[milk_sales, on = .(y4_hhid, b2)]

# add total production back in, given there is just one buyer all sales can be allocated to them
milk.s <- milkfin[,.(y4_hhid, lvstckcat = type, sold)]
milk_sales <- milk.s[milk_sales, on = .(y4_hhid, lvstckcat)]

saveRDS(milk_sales, here::here("data", "processed", "01", "buyers_milk.RDS"), compress = T)
```


## Egg buyers

```{r egg buyers}
# buyers information
produce <- readRDS(here::here("data", "processed", "01", "produce.RDS"))
# quantities
mass_eggs <- readRDS(here::here("data", "processed", "01", "mass_eggs.RDS"))

egg.s <- produce[productid == "eggs" & produced>0]
egg_sales <- egg.s[,.(y4_hhid, b1 = buyer1, b2 = buyer2)]

# merge first buyer
lf_net <- lf_network[,.(y4_hhid, b1 = lfnr_id, b1_rel = lfnr_02, b1_loc = lfnr_03)]
egg_sales <- lf_net[egg_sales, on = .(y4_hhid, b1)]

# merge second buyer
lf_net <- lf_network[,.(y4_hhid, b2 = lfnr_id, b2_rel = lfnr_02, b2_loc = lfnr_03)]
egg_sales <- lf_net[egg_sales, on = .(y4_hhid, b2)]

# add production info
eggp <- mass_eggs[,.(y4_hhid, sold)]
egg_sales <- eggp[egg_sales, on = .(y4_hhid)]

saveRDS(egg_sales, here::here("data", "processed", "01", "buyers_eggs.RDS"), compress = T)
```


## Crop buyers


### Ag network

```{r ag network hk}
ag_network <- read_dta(here::here("data", "raw", "ag_network.dta"))

ag_network <- ag_network %>% 
  clean_up() %>% 
  select(!agnr_01)

saveRDS(ag_network, here::here("data", "processed", "01", "ag_network.RDS"), compress = T)
```


```{r crop buyers}
ag_network <- readRDS(here::here("data", "processed", "01", "ag_network.RDS"))
crop_disp <- readRDS(here::here("data", "processed", "01", "crop_disp.RDS"))

# select vars
crop_disp <- crop_disp[,.(y4_hhid, type, cropid, sold, value_sale, b1, soldb1, value_b1, n_int1, b2, soldb2, value_b2, n_int2)]
crop_disp <- crop_disp[sold>0]

# merge for first buyer
ag_net <- ag_network[,.(y4_hhid, b1 = agnr_id, b1_rel = agnr_02, b1_loc = agnr_03)]
crop_sales <- ag_net[crop_disp, on=.(y4_hhid, b1)]

# merge for second buyer
ag_net <- ag_network[,.(y4_hhid, b2 = agnr_id, b2_rel = agnr_02, b2_loc = agnr_03)]
crop_sales <- ag_net[crop_sales, on=.(y4_hhid, b2)]

saveRDS(crop_sales, here::here("data", "processed", "01", "buyers_crops.RDS"), compress = T)
```

```{r tree buyers}
tree_disp <- readRDS(here::here("data", "processed", "01", "tree_disp.RDS"))

tree_disp <- tree_disp[,.(y4_hhid, type, cropid, sold, value_sale, b1, b2)]
tree_disp <- tree_disp[sold>0]

# buyers
# merge for first buyer
ag_net <- ag_network[,.(y4_hhid, b1 = agnr_id, b1_rel = agnr_02, b1_loc = agnr_03)]
tree_sales <- ag_net[tree_disp, on=.(y4_hhid, b1)]

# merge for second buyer
ag_net <- ag_network[,.(y4_hhid, b2 = agnr_id, b2_rel = agnr_02, b2_loc = agnr_03)]
tree_sales <- ag_net[tree_sales, on=.(y4_hhid, b2)]

saveRDS(tree_sales, here::here("data", "processed", "01", "buyers_trees.RDS"), compress = T)
```



