---
title: "06_Summary"
author: "Heike Rolker"
date: "2024-01-10"
output: html_document
---

```{r setup}
library(data.table)
library(tidyverse)

library(RColorBrewer)
library(plotly)
```

```{r}
# agricultural zones: https://research.csiro.au/livegaps/findings/livestock-production/dairy-production-in-tanzania/
zones <- list(
  "Lakes" = c("Kagera", "Mara", "Mwanza", "Geita", "Simiyu", "Shinyanga", "Kigoma"),
  "Coastal" = c("Mjini Magharibi", "Kaskazini Pemba", "Kaskazini Unguja", "Kusini Pemba", "Kusini Unguja", "Lindi", "Mtwara", "Morogoro", "Pwani", "Dar Es Salaam"),
  "Central" = c("Manyara", "Dodoma", "Singida", "Tabora"),
  "Northern Highlands" = c("Arusha", "Kilimanjaro", "Tanga"),
  "Southern Highlands" = c("Iringa", "Njombe", "Ruvuma", "Katavi", "Rukwa", "Mbeya"))

zones <- data.frame(region = unlist(zones), zone = rep(names(zones), lengths(zones)))

zones <- setDT(zones) 
```

Overview:

Final *general* files that can serve as input for any subsequent chapter building on LSMS data. 

```{r functions}
clear.labels <- function(x) {
  if(is.list(x)) {
    for(i in seq_along(x)) {
      class(x[[i]]) <- setdiff(class(x[[i]]), 'labelled')
      attr(x[[i]],"label") <- NULL
    }
  } else {
    class(x) <- setdiff(class(x), "labelled")
    attr(x, "label") <- NULL
  }
  return(x)
}  


```

# 1) All production & destination (overview and mass-balance files)

Overview:

- **_long** files list key information for each item produced by households (produced, sold, consumed), ag/milk produce in seprate list

- **mfa_** balanced quantities as input for mfa

Actions:

- bring animals and hides together - x

- bring crops and produce together - x

- add residue and yield gaps to crops - **TBD**

- long format of core info: production, sales, consumption, all destinations, uncertain; merge with FCT info - x [add food groups]

- **missing mass balance estimates for each file** - 


```{r}
mass_allcrops <- readRDS("2_data/processed/mass_allcrops.RDS") %>% 
  clear.labels() # no code exists
mass_agprod <- readRDS("2_data/processed/mass_agprod_new.RDS") %>% 
  clear.labels()
# mass_residue
# mass_yg

agprod_long <- readRDS("2_data/processed/mass_agprod_long_new.RDS") %>%  # details of agricultural produce
  clear.labels()

mass_animals <- readRDS("2_data/processed/animals_mass.RDS") %>% 
  clear.labels() # alternative is remove_labels{crosstable}

mass_hides <- readRDS("2_data/processed/mass_hides.RDS") %>% 
  clear.labels()

mass_milk <- readRDS("2_data/processed/mass_milk.RDS") %>% 
  clear.labels()

mass_eggs <- readRDS("2_data/processed/mass_eggs_fin.RDS") %>% 
  clear.labels()

```

# 3) All exclusions

excl_ files
```{r}

# save all as csv
excl_agprod <- read_csv("2_data/processed/excl_agprod.csv")
excl_allcrops <- read_csv("2_data/processed/excl_allcrops.csv")
excl_animals <- read_csv("2_data/processed/excl_animals.csv")
excl_crops <- read_csv("2_data/processed/excl_crops.csv")
excl_trees <- read_csv("2_data/processed/excl_trees.csv")
excl_milk <- read_csv("2_data/processed/excl_milk.csv")
excl_eggs <- read_csv("2_data/processed/excl_eggs.csv")
excl_long <- read_csv("2_data/processed/excl_long.csv")

excl_list <- rbind(excl_agprod, excl_allcrops, excl_animals, excl_crops, excl_trees, excl_milk, excl_eggs, excl_long) 
# detailed list of exclusions
write.csv(excl_list, "2_data/final/excl_list.csv")

excl_final <- excl_list %>% 
  filter(!is.na(excl)) %>% 
  select(y4_hhid) %>% 
  unique()

write.csv(excl_final, "2_data/final/excl_3a.csv")

# other chapters
write.csv(excl_final, "~/Library/CloudStorage/OneDrive-UniversityofBristol/03b_MFA_connect/2_data/final/excl_3a.csv")
```

Mass balancing used for MFA


## 1a) Mass Balancing


```{r}
# applicable to all
calc <- function(df) df[, lapply(.SD, sm), .SDcols=is.numeric, by=.(type)] # change function and grouping as required; for 3c this could stay item; use groupsgroups for other aggregation levels
# col sum
cl <- function(df) df[, lapply(.SD, sm), .SDcols=is.numeric] # change function as required

sm <- function(x) sum(x, na.rm = TRUE)
rd <- function(df) df %>% dplyr::mutate_if(is.numeric, round, 1)
md <- function(x) median(x, na.rm=TRUE)
mn <- function(x) mean(x, na.rm=TRUE)
```


### Crops
```{r}
# Crops for MFA - 
crops <- mass_allcrops %>% 
  dplyr::rename(produced = harvest, item = cropid) %>% 
  left_join(select(mass_agprod, y4_hhid, item = cropid, processing = input, prodproduced = produced, inprodsold = input_sales, prodsold = sold, inprodcons = input_cons, prodconsumed = consumed, waste = uncertain), by = c("item", "y4_hhid")) %>% 
  mutate_if(is.numeric, ~replace(., is.na(.), 0)) %>%
  mutate(uncertain = produced - smd, # ignores the issue resulting from processing being unclear
         uncertain2 = uncertain - processing, # includes processing; many negative values..
         missing = ifelse(uncertain<0, uncertain * -1, 0),
         unallocated = ifelse(uncertain>0, uncertain, 0)) # if uncertain an excess value then allocated to "missing quantities" 

saveRDS(crops, "2_data/final/mfa_crops.RDS", compress = T)
```


### Animals

```{r}
# Animals for MFA
animals <- mass_animals %>% 
  # filter(slaughter>0) %>% 
  select(!ew) %>%  # no need
  mutate(
    consumed = meat + offal,
    uncertain = total_weight - sold_weight - meat - offal - hides - inedible, # should be 0 for all of them... difference from breakdown not fully matched up, use difference as measure of uncertainty
    missing = ifelse(uncertain<0, uncertain * -1, 0), # more allocation than production
    unallocated = ifelse(uncertain>0, uncertain, 0), # assumes allocation underestimated (no waste)
    upper = total_weight + missing, # upper total weight estimate
    lower = total_weight - unallocated, # lower total weight estimate
    mean = (upper + lower + total_weight)/3 # mean weight estimate
  )

saveRDS(animals, "2_data/final/mfa_animals.RDS", compress = T)

hides <- mass_hides

saveRDS(hides, "2_data/final/mfa_hides.RDS", compress = T)
```


### Milk
```{r}
# Milk for MFA
milk <- mass_milk %>% 
  mutate(
    smd = consumed + sold + processed, # here processed sales are not included in the raw sales (at least not necessarily)
    uncertain = milk-smd,
    missing = ifelse(uncertain<0, uncertain * -1, 0), # more allocation than production
    unallocated = ifelse(uncertain>0, uncertain, 0), # less allocation than produciton
    item = "milk"
  ) %>% 
  dplyr::rename(produced = milk) # uses mostly quantities sold as processing inputs, assumption: data checking error.
  
saveRDS(milk, "2_data/final/mfa_milk.RDS", compress = T)
```


### Eggs
```{r}
# Eggs for MFA
eggs <- mass_eggs %>% 
  select(y4_hhid, item = productid, feed, grazed, produced, sold, consumed, uncertain = missing) %>%  # type to be replaced by FCT food group
  mutate(
    # there should be a range of consumption estimates...
    smd = sold + consumed, # have it in other dts, necessary?
    uncertain = produced - smd, # to make sure it's correct
    missing = ifelse(uncertain<0, uncertain * -1, 0), # more allocation than production
    unallocated = ifelse(uncertain>0, uncertain, 0) # less allocation than production
  ) %>% 
  add_column(
    type = "Eggs" # consistent with groupsgroups?
  )

saveRDS(eggs, "2_data/final/mfa_eggs.RDS", compress = T)

```


## 1b) Long file

```{r crops}
# Listing crops - x
# crops
crops_long <- crops %>% 
  select(y4_hhid, type, item, produced, sold, consumed, smd, uncertain, missing, unallocated) %>% 
  # groupsgroups
  mutate(
    product_type = ifelse(type == "cashcrops", "Other crops", "Food crops"),
    subtype = type) %>%  # required twice
  add_column(
    category = "Plant"
  ) %>% 
  select(y4_hhid, category, product_type, item, type, subtype, produced, sold, consumed, smd, uncertain, missing, unallocated)

# agricultural products
crops_long1 <- agprod_long %>% 
  add_column(
    category = "Plant",
    product_type = "Processed crop"
  )  %>% 
  left_join(crops_list, by = "cropid") %>% 
  dplyr::rename(
    group = cropid,
    subtype = product
  ) %>% 
  mutate(produced = as.numeric(produced),
         sold = as.numeric(sold),
         consumed = as.numeric(consumed)) %>% 
  select(y4_hhid, category, product_type, item, type, subtype, produced, sold, consumed)
```


Keep animal info and hides as separate dataframes as not to lose detail (ie livestock).


```{r animals}
# Listing animals - x
animals_fin <- readRDS("2_data/processed/animals_fin.RDS")
# Total slaughter product, and weight allocated to destinations - x
animals_long <- animals_fin %>% 
  clear.labels() %>% 
  filter(slaughter>0) %>% 
  mutate(consumed = sl_weight - sold_weight) %>% 
  select(y4_hhid, type, item = lvstckid, produced = sl_weight, sold = sold_weight, consumed = cons_weight) %>%  
  # add grouping
  add_column(
    category = "Animal",
    product_type = "Slaughter product",
    subtype = "other"
  ) %>% 
  select(y4_hhid, category, product_type, item, type, subtype, produced, sold, consumed)

# Meat and offal produced from animals slaughtered (consumption only) - x
animals_long1 <- animals %>%
# list meat & offal as edible items in long data frame format
  melt(id.vars=c("y4_hhid", "lvstckid", "type"),
    measure.vars=c("meat", "offal"), #  meat and offal are consumed
    variable.name="item",
    value.name="consumed") %>% 
  mutate(item = paste(lvstckid, item, sep = " "), 
         subtype = lvstckid) %>% 
  # detail based on groupsgrousp
  add_column(
    category = "Animal",
    product_type = "Animal product"
  ) %>% 
  mutate(# item = paste(subtype, item, sep = " "), # to stay consistent with mapping against fct,  duplication here
        produced = consumed, 
        sold = 0) %>% 
  mutate(produced = as.numeric(produced),
         sold = as.numeric(sold),
         consumed = as.numeric(consumed)) %>% 
  select(y4_hhid, category, product_type, item, type, subtype, produced, sold, consumed)

# Hides - x
hides_long <- mass_animals %>% 
  filter(slaughter>0) %>% 
  group_by(y4_hhid, type) %>% 
  summarise_if(is.numeric, sum) %>% 
  left_join(select(mass_hides, y4_hhid, type, prodhides = pprod, hidescons = missing, hidessold = sold2), by = c("y4_hhid", "type")) %>% 
  filter(!is.na(prodhides)) %>% 
  mutate(
    item = paste("hides", type, sep = "-")
  ) %>% 
  select(y4_hhid, type, item, hides, produced = prodhides, sold = hidessold, consumed = hidescons) %>%  # hides estimate larger than hides produced in some cases, would require to review the data again to reconcile -> ignore, too time consuming, too little pay-off
  mutate(uncertain = hides - produced) %>% 
  select(!hides) %>% 
  add_column(
    category = "Animal",
    product_type = "other",
    subtype = "other"
  ) %>% 
  select(y4_hhid, category, product_type, item, type, subtype, produced, sold, consumed, uncertain)
```

```{r milk}
# Listing milk
# Milk - x
milk_long <- milk %>% 
  mutate(sold = sold+psold, # assumes raw sold and processed sold are different accounts
         item = paste(item, type, sep = " - ")) %>% 
  add_column(
    category = "Animal",
    product_type = "Animal product",
    subtype = "Milk"
  ) %>% 
  select(y4_hhid, category, product_type, item, type, subtype, produced, sold, consumed, smd, uncertain, missing, unallocated)
  
# Processed milk - x
milk_long1 <- milk %>% 
  mutate(
    item = paste("Processed milk", type, sep = " - "),
    pcons = processed - psold # in a small number of cases this works, assumes no waste but all sold or consumed
  ) %>% 
  add_column(
    category = "Animal",
    product_type = "Animal product",
    subtype = "Processed milk"
  ) %>% 
  select(y4_hhid, category, product_type, item, type, subtype, produced = processed, sold = psold, consumed = pcons, smd, uncertain, missing, unallocated)
```


```{r eggs}
# Listing eggs - x
eggs_long <- eggs %>% 
  add_column(
    category = "Animal",
    product_type = "Animal product",
    subtype = "Eggs",
    type = "Eggs"
  ) %>% 
  select(y4_hhid, category, product_type, item, type, subtype, produced, sold, consumed, smd, uncertain, missing, unallocated)
```

- all production in one dt: y4_hhid, item, (type from FCT), produced, sold, consumed (ignores ag produce but may have to account for processed milk), smd (includes all other destinations), uncertain

```{r}

# FINAL LIST 
produce_long <- crops_long %>% 
  # bind all product lists
  dplyr::bind_rows(animals_long1, animals_long, crops_long1, hides_long, milk_long, milk_long1, eggs_long) %>% 
  # check no duplications -> just a few
  group_by(y4_hhid, category, product_type, item, type, subtype) %>% 
  summarise(across(is.numeric, ~ sum(.x, na.rm = TRUE))) %>%  # code deprecated, should be replaced
  ungroup() %>% 
  mutate(across(item:subtype, ~str_to_sentence(.x))) %>% 
  # adjust order
  select(y4_hhid, category, product_type, type, subtype, item, produced, sold, consumed, smd, uncertain)

saveRDS(produce_long, "2_data/final/produce_long.RDS", compress = T)

# Group file
groupsgrps <- produce_long %>% 
  select(category, product_type, type, subtype, item) %>% 
  unique()

write_csv(groupsgrps, "2_data/reference/groupsgrps.csv")

# save to other chapter where needed
saveRDS(produce_long, "~/Library/CloudStorage/OneDrive-UniversityofBristol/03b_MFA_connect/2_data/files/produce_long.RDS")
write_csv(groupsgrps, "~/Library/CloudStorage/OneDrive-UniversityofBristol/03_LSMS/TZAwave4/files/groupsgrps.csv") # update location
```

```{r produce long exclusions}
excl_long <- produce_long %>% group_by(y4_hhid) %>% 
  summarise(sum = sm(produced)) %>% 
  filter(sum == 0) %>% 
  add_column(
    item = "Produce long",
    excl = "No production"
  ) %>% 
  select(!sum)

write.csv(excl_long, "2_data/processed/excl_long.csv")
```

### Find all types not produced by household

```{r}
# list all types from groupsgroups
# merge & identify missing
```


### Link to FCT

```{r}
# add FCT info : here:here check if merged fully (some names might have been changed along the way...)
LSMS_AG_FCT <- read_csv("2_data/reference/AG_FCT_final.csv") %>% 
  select(item = product, group, Edible_CF:Cholesterol_CHOLE_mg) %>% 
  unique() # contains some duplication

produce_long_fct <- produce_long %>% 
  left_join(LSMS_AG_FCT, by = "item")
# give group to cashgroups


saveRDS(produce_long_fct, "2_data/final/produce_long_fct.RDS", compress = T)
```


## 1b) All primary buyers

```{r merge all primary buyers, eval=FALSE}
# read all data
buyers_animals <- readRDS("2_data/processed/buyers_animals.RDS")
buyers_crops <- readRDS("2_data/processed/buyers_crops.RDS")
buyers_eggs <- readRDS("2_data/processed/buyers_eggs.RDS")
buyers_milk <- readRDS("2_data/processed/buyers_milk.RDS")
buyers_trees <- readRDS("2_data/processed/buyers_trees.RDS")

# crop buyer 
crop_sales <- readRDS("2_data/processed/crop_sales.RDS") # contains more detail

# calculate share for buyers
crop_sales[, `:=` (b1_perc = soldb1/sold*100,
                   b2_perc = soldb2/sold*100)]

# factor variables
crop_sales[, `:=` (b1_rel = as.factor(b1_rel),
                   b1_loc = as.factor(b1_loc),
                   b2_rel = as.factor(b2_rel),
                   b2_loc = as.factor(b2_loc),
                   type = as.factor(type))] #should be...

# overview
# table(crop_sales$type, crop_sales$b1_rel)
# table(crop_sales$type, crop_sales$b1_loc)

b1_crops <- crop_sales[,.(y4_hhid, product = type, type = cropid, sold = soldb1, b1, b1_rel, b1_loc)]

# tree buyers
b1_trees <- buyers_trees[,.(y4_hhid, product = type, type = cropid, sold, b1, b1_rel, b1_loc)]

# animal buyers
b1_animals <- buyers_animals[,.(y4_hhid, type, product = lvstckid, sold = sold_weight, b1, b1_rel, b1_loc)]

# milk buyers
buyers_milk[, product := "milk"]
b1_milk <- buyers_milk[,.(y4_hhid, type = lvstckcat, product, sold, b1, b1_rel, b1_loc)]

# egg buyers
buyers_eggs[,`:=` (type = "animal product",
                 product = "eggs")]

b1_eggs <- buyers_eggs[,.(y4_hhid, type, product, sold, b1, b1_rel, b1_loc)]

# merge all buyers
b1 <- list(b1_animals, b1_eggs, b1_milk, b1_crops, b1_trees)

#b1 <- lapply(b1, function(df){df <- df[, `:=` (b1_rel = replace_na(b1_rel, "unknown"), b1_loc = replace_na(b1_loc, "unknown"))]})
# overcome class errors
b1 <- lapply(b1, function(df){df <- df[, `:=` (b1 = as.factor(b1),
                                               b1_rel = as.factor(b1_rel),
                                               b1_loc = as.factor(b1_loc),
                                               type = as.factor(type),
                                               product = as.character(product),
                                               y4_hhid = as.character(y4_hhid),
                                               sold = as.numeric(sold))]})

# merge buyers into complete list
names(b1) <- c("animals", "eggs", "milk", "crops", "trees")

b1 <- do.call("rbind", lapply(names(b1), function(x) cbind(index = x, b1[[x]])))

b1 <- b1[sold>0] # exclude where no sales

# match to groups groups pattern
b1_gr <- b1 %>% 
  mutate(
    type = as.character(type),
    product = as.character(product),
    # category
    category = fcase(
      index == "trees" | index == "crops", "Plant",
      index == "eggs" | index == "milk" | index == "animals", "Animal"
    ),
    # product type
    product_type = fcase(
      index == "animals", "Slaughter product",
      product == "cashcrops", "Other crops",
      index == "eggs" | index == "milk", "Animal products"
    ),
    product_type = ifelse(is.na(product_type), "Food crops", product_type),
    # subtype
    subtype = fcase(
      product == "milk", "Milk", 
      product == "eggs", "Eggs",
      product_type == "Slaughter product", "other"),
    subtype = ifelse(is.na(subtype), product, subtype),
    # type
    type_new = fcase(
      product == "eggs", "Poultry",
      product == "milk", product,
      index == "animals", type
    ),
    type_new = ifelse(is.na(type_new), product, type_new),
    # item
    item = fcase(
      product == "milk", paste("milk", type, sep = " - "),
      index == "animals", product),
    item = ifelse(is.na(item), type, item)
    ) %>% 
  select(
    y4_hhid, category, product_type, type_new, subtype, item, sold, b1, b1_rel, b1_loc) %>% 
  mutate(
    new_name = fcase(
      is.na(b1_rel), "Unknown",
      b1_rel == "neighbor" | b1_rel == "friend" | b1_rel == "relative", "Family, friends & neighbours",
      b1_rel == "market" | b1_rel == "open market" | b1_rel == "grocery/local merchant", "Formal markets",
      b1_rel == "cooperative union"| b1_rel == "distribution officer" | b1_rel == "abbatori/factory" | b1_rel == "farmers party", "Formal trader or processor",
      b1_rel == "employer" | b1_rel == "business contact"| b1_rel == "bank (commercial)" | b1_rel == "religious institution", "Institutions & commerical buyers",
      b1_rel == "private business person", "Informal market & vendors"
      # private business person:This is usually a vendor or a hawker who sells goods by walking door to door – like cigarette/water sellers or vegetable vendors in Dar.
    ),
    new_location = fcase(
      is.na(b1_loc), "Unknown",
      b1_loc == "within the village" | b1_loc == "near the village", "In village proximity",
      b1_loc == "other region", "Other region", # keep until I come up with something better...
      b1_loc == "other district", "Other district",
      b1_loc == "near the town", "In town proximity",
      b1_loc == "across the border", "Across the border"
    )) %>%
  mutate(across(is.character, ~str_to_sentence(.x)))

saveRDS(b1_gr, "2_data/final/b1.RDS", compress = T)
```
```{r export for python}
# Export RDS inputs to CSV for Python translation
library(here)
saveRDS_to_csv <- function(rds_name, csv_name) {
  readRDS(here::here("data", "processed", "01", rds_name)) |>
    data.table::fwrite(here::here("data", "processed", "01", csv_name))
}

saveRDS_to_csv("mfa_crops.RDS",   "mfa_crops.csv")
saveRDS_to_csv("mfa_eggs.RDS",    "mfa_eggs.csv")
saveRDS_to_csv("mfa_milk.RDS",    "mfa_milk.csv")
saveRDS_to_csv("mfa_animals.RDS", "mfa_animals.csv")
```




