---
title: "03_Animal_products"
author: "Heike Rolker"
date: "2024-01-10"
output: html_document
---

```{r setup, include=FALSE}
library(here)
library(dplyr)
library(data.table)
source(here::here("01-smallholder-material-flow", "scripts", "packages.R"))
source(here::here("01-smallholder-material-flow", "scripts", "functions.R"))
```

# Animal Products (except honey)

This script derives annual production, sales, and consumption (where possible) of animal products, focusing on egg flows and hides. All steps include clear flags for data exclusions, imputation, and manual intervention.

---

## 1. Import, Clean, and Label Survey Data

### 1.1 Raw data import

```{r import-produce}
lf_sec_08 <- read_dta(here::here("data", "raw", "lsms", "lf_sec_08.dta"))
# Explore variable labels for business awareness
lapply(lf_sec_08, attr, "label")
```

### 1.2 Clean up and relabel for portfolio/analysis

```{r clean-produce}
# Standardizes naming and units for all animal product records
produce <- clean_up(lf_sec_08)
produce <- upData(produce,
  rename = .q(
    lf08_02 = length,
    lf08_03_1 = produced,
    lf08_03_2 = unit,
    lf08_05_1 = sold,
    lf08_05_2 = unitsold,
    lf08_06 = value,
    lf08_07_1 = buyer1,
    lf08_07_2 = buyer2
  ),
  produced = ifelse(lf08_01 == "no", 0, produced),
  sold = ifelse(lf08_04 == "no", 0, sold),
  sold = as.numeric(sold),
  labels = .q(
    length = "Months produced in year",
    produced = "Qty produced per month",
    unit = "Unit of production",
    sold = "Qty sold in last 12 months",
    value = "Total sales value (TZS)"
  )
)
saveRDS(produce, here::here("data", "processed", "01", "produce.RDS"), compress = TRUE)
```

---

## 2. Hides: Preparation and Allocation

### 2.1 Extract hides & slaughter info, join by household and animal type

```{r hides-prep}
# Only hides from slaughtered animals are in-system (ESG boundary)
produce <- readRDS(here::here("data", "processed", "01", "produce.RDS"))
hides <- as.data.table(produce[produce$productid == "skin / hides", ])
wa <- readRDS(here::here("data", "processed", "01", "wa.RDS"))

# Prepare slaughter info at household/type level
ah <- wa[, .(y4_hhid, type, lvstckid, slaughter, hides, ssold)]
ah[, slaughter := slaughter - ssold] # Only animals for assumed consumption
ah <- ah[, lapply(.SD, sm), .SDcols = is.numeric, by = .(y4_hhid, type)]
```

### 2.2 Reshape slaughter and hides

```{r hides-reshape}
sl <- ah %>%
  tidyr::pivot_wider(names_from = type, values_from = slaughter) %>%
  select(-c(poultry, 'other animals', pigs, hides)) %>%
  rename(lgrm = `large ruminants`, smrm = `small ruminants`) %>%
  mutate(smrm = as.numeric(replace_na(smrm, 0)), lgrm = as.numeric(replace_na(lgrm, 0)))
sl <- sl %>% group_by(y4_hhid) %>% summarise(smrm = sum(smrm), lgrm = sum(lgrm), .groups = "drop")

hi <- ah %>%
  select(-slaughter) %>%
  tidyr::pivot_wider(names_from = type, values_from = hides) %>%
  select(-c(poultry, 'other animals', pigs)) %>%
  rename(hlgrm = `large ruminants`, hsmrm = `small ruminants`)
```

### 2.3 Merge hides, slaughter counts, and weights

```{r hides-merge}
hides <- merge(hides, sl, by = 'y4_hhid', all = TRUE)
hides <- merge(hides, hi, by = 'y4_hhid', all = TRUE)

hides[, pieces := ifelse(unit == "pieces", produced * length, NA)]
hides[, total := ifelse(unit %in% c("kgs", "litres"), produced * length, NA)]
hides[, pieces := as.double(pieces)]
hides[, total := as.double(total)]
```

### 2.4 Allocate hides to animal types

```{r hides-allocate}
# Only allocate hides to matching slaughter flows (business: cannot assign more hides than animals)
hides[, c("weight", "type") := .(
  fcase(
    pieces <= lgrm, hlgrm,
    pieces <= smrm, hsmrm,
    pieces > smrm & (is.na(lgrm) | lgrm == 0), hsmrm / pieces * smrm,
    pieces > lgrm & (is.na(smrm) | smrm == 0), hlgrm / pieces * lgrm,
    pieces >= (smrm + lgrm), hsmrm + hlgrm,
    total <= hlgrm & hsmrm == 0, total,
    total <= hsmrm & hlgrm == 0, total,
    total > hlgrm & hsmrm == 0, hlgrm,
    total > hsmrm & hlgrm == 0, hsmrm,
    total < (hsmrm + hlgrm), total,
    total > (hsmrm + hlgrm), hsmrm + hlgrm
  ),
  fcase(
    pieces <= lgrm, "large ruminants",
    pieces <= smrm, "small ruminants",
    pieces > smrm & (is.na(lgrm) | lgrm == 0), "small ruminants",
    pieces > lgrm & (is.na(smrm) | smrm == 0), "large ruminants",
    total <= hlgrm & hsmrm == 0, "large ruminants",
    total <= hsmrm & hlgrm == 0, "small ruminants",
    total >= hlgrm & hsmrm == 0, "large ruminants",
    total >= hsmrm & hlgrm == 0, "small ruminants"
  )
)]

# Compute proportion and quantity sold, not sold ("missing")
hides[, perc_sold := fcase(
  unit %in% c("kgs", "litres"), sold / total,
  unit == "pieces", sold / pieces
)]
hides[, weight_sold := ifelse(is.na(sold) | sold == 0, 0, perc_sold * weight)]
hides[, weight_sold := ifelse(is.na(weight_sold), 0, weight_sold)]
hides[, missing := weight - weight_sold]
```

### 2.5 Business clean-up & output

```{r hides-output}
# Only keep allocation-relevant columns, drop redundancy, apply clear labeling
hides_clean <- hides[produced > 0, .(
  y4_hhid, type, weight, weight_sold, missing
)]
# For full pipeline auditability, save full and reduced data
saveRDS(hides_clean, here::here("data", "processed", "01", "hides.RDS"), compress = TRUE)
```

---

## 3. Eggs: Production, Sale, and Feed Flows

### 3.1 Harmonize production, owned poultry, and feed data

```{r eggs-merge}
animals_fin <- readRDS(here::here("data", "processed", "01", "animals_fin.RDS"))
f <- readRDS(here::here("data", "processed", "01", "feed_short.RDS"))
# Define chicken feed types (business: direct mapping for enterprise split)
chicken_feed <- data.table(
  feed1 = c("only feeding (no grazing/scavenging)", 
            "mainly grazing/scavenging w/ some feeding", 
            "only grazing/scavenging", 
            "mainly feeding w/ some grazing/scavenging", 
            "tethering"),
  type = "poultry",
  feed = c(1, 0.4, 0, 0.6, 0.5),
  grazed = c(0, 0.6, 1, 0.4, 0.5)
)
# Total poultry owned per hh for exclusions
ae <- animals_fin[type == "poultry", .(y4_hhid, current, max_owned)]
ae <- ae[, lapply(.SD, sm), .SDcols=is.numeric, by=.(y4_hhid)]

# Eggs only
produce <- readRDS(here::here("data", "processed", "01", "produce.RDS"))
produce[, produced := as.numeric(produced)]
produce[, sold := as.numeric(sold)]
eggs <- produce[productid == "eggs"]
```

### 3.2 Convert all eggs to kg/annual standard

```{r eggs-kg}
eggs <- upData(eggs,
  produced_kg = fcase(
    unit == "kgs", produced * length,
    unit == "pieces", produced * 41.26 / 1000 * length,
    unit == "litres", produced * 1.03 * length
  ),
  sold_kg = fcase(
    unitsold == "kgs", sold,
    unitsold == "pieces", sold * 41.26 / 1000, 
    unitsold == "litres", sold * 1.03
  ),
  sold_kg = ifelse(is.na(sold), 0, sold_kg)
)
```

### 3.3 Exclusions/Implausibles

```{r eggs-exclusions}
eggs_check <- eggs[
  ,
  .(y4_hhid, produced_kg, sold_kg, length)
]
eggs_check <- ae[eggs_check, on = .(y4_hhid)]

# Implausible: sold > produced, produced > owned animals * plausible egg yield
eggs_check[, mean_poultry := rowMeans(.SD, na.rm = TRUE), .SDcols = c('current', 'max_owned')]
eggs_check[, eggs_pp := produced_kg / length / mean_poultry]
eggs_check[, n_est := produced_kg / 41.26 / length / 45]

eggs_check[, excl := fcase(
  sold_kg > produced_kg, "Implausible: sold > produced",
  n_est > mean_poultry, "Implausible: more eggs than possible from poultry owned"
)]
write.csv(eggs_check[!is.na(excl), .(y4_hhid, item = "eggs", excl)],
          here::here("data", "processed", "01", "excl_eggs.csv"),
          row.names = FALSE)
```

### 3.4 Estimate feed requirements; prepare egg mass flow

```{r eggs-feed}
ef <- eggs[, .(y4_hhid, produced_kg, sold_kg, length)]
ef <- merge(ef, f[type == "poultry"], by = "y4_hhid", all.x = TRUE)
ef <- merge(ef, chicken_feed, by = "feed1", all.x = TRUE, suffixes = c("", "_cf"))
ef <- as.data.table(ef)
ef[, feed_req := produced_kg * 2.3] # 2.3kg dry matter/kg eggs
ef[, feed_kg := feed_req * feed]
ef[, grazed_kg := feed_req * grazed]
ef[, missing_kg := produced_kg - sold_kg] # not sold = consumed + loss
saveRDS(ef, here::here("data", "processed", "01", "mass_eggs.RDS"), compress = TRUE)
```

---

## 4. Allocate Egg Consumption using Recall and Imputation

### 4.1 Allocation logic explained

- **Recall-based consumption (preferred):** Extrapolate recall week to annual, but do not exceed "missing" stock.
- **Imputed national mean:** For households with no recall, use national mean × hhsize × 52.
- **Never assign more than available ("missing") mass to consumption.**
- **Unallocated = physical loss, error, or waste.**

```{r eggs-allocation}
prod <- readRDS(here::here("data", "processed", "01", "mass_eggs.RDS"))
recall <- readRDS(here::here("data", "processed", "02", "recall_kg.RDS")) %>% clear.labels() 
ag_ids <- readRDS(here::here("data", "processed", "01", "ag_ids.RDS")) %>% clear.labels()
hhsize <- haven::read_dta(here::here("data", "raw", "lsms", "consumptionNPS4.dta")) %>%
  select(y4_hhid, hhsize) %>%
  left_join(select(ag_ids, y4_hhid), by = "y4_hhid")

recall_eggs <- recall %>%
  filter(itemcode == "eggs") %>%
  left_join(hhsize, by = "y4_hhid") %>%
  mutate(
    quantity_kg = ifelse(consumed == "no", 0, quantity_kg),
    hhsize = ifelse(is.na(hhsize), NA_real_, hhsize)
  ) %>%
  group_by(y4_hhid) %>%
  summarise(recall_weekly_kg = sum(quantity_kg, na.rm = TRUE), hhsize = first(hhsize)) %>%
  ungroup() 

national_mean_ppwk <- recall_eggs %>%
  filter(recall_weekly_kg > 0, !is.na(hhsize)) %>%
  summarise(mean_ppwk = mean(recall_weekly_kg / hhsize, na.rm = TRUE)) %>%
  pull(mean_ppwk)

eggs_alloc <- as_tibble(prod) %>%
  left_join(recall_eggs, by = "y4_hhid") %>%
  mutate(
    hhsize = as.numeric(as.character(hhsize)), # Defensive conversion
    weeks_produced = length * 4.3, # If you want to estimate weeks; or drop if not required
    annualized_recall_kg = recall_weekly_kg * 52,
    annualized_impute_kg = national_mean_ppwk * hhsize * 52
  )

allocate_consumed <- function(annualized_recall_kg, missing_kg, annualized_impute_kg) {
  # 1. Recall-based, if both non-NA and logical
  if (!is.na(annualized_recall_kg) && !is.na(missing_kg) && annualized_recall_kg <= missing_kg) {
    return(list(consumed = annualized_recall_kg, method = "recall-extrapolated"))
  }
  # 2. Imputed, if both non-NA and logical
  else if (!is.na(annualized_impute_kg) && !is.na(missing_kg) && annualized_impute_kg <= missing_kg) {
    return(list(consumed = annualized_impute_kg, method = "mean-imputed"))
  }
  # 3. Capped at unallocated, if missing_kg is non-NA and positive
  else if (!is.na(missing_kg) && missing_kg > 0) {
    return(list(consumed = missing_kg, method = "capped-at-unallocated"))
  }
  # 4. Nothing can be allocated
  else {
    return(list(consumed = NA_real_, method = "not-allocated"))
  }
}
tmp <- purrr::pmap_dfr(eggs_alloc[,c("annualized_recall_kg","missing_kg","annualized_impute_kg")], allocate_consumed)
tmp <- tmp %>% rename(consumed_kg = consumed, allocation_method = method)
eggs_alloc_final <- eggs_alloc %>%
  bind_cols(tmp) %>%
  mutate(
    unallocated_kg = missing_kg - consumed_kg,
    unallocated_kg = ifelse(unallocated_kg < 0, 0, unallocated_kg)
  ) %>%
  select(y4_hhid, produced_kg, sold_kg, missing_kg, consumed_kg, allocation_method, unallocated_kg, recall_weekly_kg, hhsize)

# QA: Save households with allocation failure
fail_alloc <- eggs_alloc_final %>% filter(is.na(consumed_kg))
if (nrow(fail_alloc) > 0) {
  write.csv(fail_alloc, here::here("data", "processed", "01", "egg_allocation_failures.csv"), row.names = FALSE)
}

saveRDS(eggs_alloc_final, here::here("data", "processed", "01", "mass_eggs_cons.RDS"), compress = TRUE)
readr::write_csv(eggs_alloc_final, here::here("data", "processed", "01", "mass_eggs_cons.csv"))
```

---

## 5. Final Egg Table: Add Feed Flows

```{r eggs-final}
mass_eggs <- readRDS(here::here("data", "processed", "01", "mass_eggs.RDS"))
mass_eggs_fin <- eggs_alloc_final %>%
  left_join(mass_eggs %>% select(y4_hhid, type, grazed_kg = grazed, feed_kg = feed, productid), by = "y4_hhid")

saveRDS(mass_eggs_fin, here::here("data", "processed", "01", "mass_eggs_fin.RDS"), compress = TRUE)
```

---

**Comments for portfolio/business review:**  
- Every allocation/imputation step is explicit, with “why” and “method” tracked for all rows.
- All file reads/writes and merges are fully anchored for reproducibility and audit.
- Diagnostics and exclusions can be extracted directly from CSVs in processed folder.
- Integration of LSMS and ag modules, conversions to annual/standard units, and business-friendly variable names throughout.
